---
title: "SAIL data extract"
execute:
  echo: true
  warning: false
  message: false
---

```{r cache_config, include=FALSE}
use_cache <- TRUE
cache_dir  <- file.path(getwd(), "cache_desc")
dir.create(cache_dir, showWarnings = FALSE, recursive = TRUE)

cache_or_run <- function(cache_file, expr) {
  if (isTRUE(use_cache) && file.exists(cache_file)) {
    message("[cache] Loading from disk: ", basename(cache_file))
    readRDS(cache_file)
  } else {
    message("[cache] Running computation: ", basename(cache_file))
    result <- expr()
    saveRDS(result, cache_file)
    message("[cache] Saved: ", basename(cache_file))
    result
  }
}
```

```{r data_source, include=FALSE}

data_source            <- "sail"
react_model_file <- "P:/kingd/GitKraken/1483_Paper-3/Re-run - New classes/msm_cvd_object_FINAL_EXCL_NHDL.qs"

# ── Mapping helpers ────────────────────────────────────────────────────────────
map_ehr_ethnicity <- function(x) {
  dplyr::case_when(
    grepl("white",       x, ignore.case=TRUE) ~ "White",
    grepl("south asian|indian|pakistani|bangladeshi", x, ignore.case=TRUE) ~ "South Asian",
    grepl("black|african|caribbean", x, ignore.case=TRUE) ~ "Black",
    grepl("mixed",       x, ignore.case=TRUE) ~ "Mixed",
    grepl("chinese|other asian|other",    x, ignore.case=TRUE) ~ "Other",
    TRUE                                      ~ "White"
  )
}

map_ehr_smoking <- function(x) {
  x <- as.character(x)
  dplyr::case_when(
    x %in% c("0","Never","Non-smoker","N")            ~ "Non-smoker",
    x %in% c("1","Ex","Ex-smoker","Former","E")       ~ "Ex-smoker",
    x %in% c("2","Light","1-9","<10")                 ~ "Light",
    x %in% c("3","Moderate","10-19","10-20")          ~ "Moderate",
    x %in% c("4","Heavy","20+",">20",">=20")          ~ "Heavy",
    TRUE                                               ~ "Non-smoker"
  )
}

map_smoke_cat <- function(s) {
  mp <- c("Non-smoker"=0L, "Ex-smoker"=1L, "Light"=2L, "Moderate"=3L, "Heavy"=4L)
  as.integer(mp[s])
}

map_ethrisk <- function(e) {
  mp <- c("White"=1L, "South Asian"=3L, "Black"=6L, "Mixed"=2L, "Other"=9L)
  v  <- mp[e]; v[is.na(v)] <- 1L; as.integer(v)
}

clamp_vec <- function(x, lo, hi) pmin(pmax(x, lo), hi)
```

```{r libs, include=FALSE}
suppressPackageStartupMessages({
  library(ggplot2)
  library(dplyr)
  library(tidyr)
  library(patchwork)
  library(ggh4x)
  library(knitr)
  library(kableExtra)
  library(scales)
  library(plotly)
  library(mice)
  library(qs)
})
dir.create("figures", showWarnings = FALSE)
```

```{r qrisk_functions, include=FALSE}
# ── QRISK2-2015 female ────────────────────────────────────────────────────────
qrisk2_2015_female <- function(age, bmi, sbp, rati, town, smoke_cat, ethrisk,
                                b_AF=0, b_ra=0, b_renal=0, b_treatedhyp=0,
                                b_type1=0, b_type2=0, fh_cvd=0, surv=10L) {
  survivor <- c(0, 0.999128758907318, 0.998222172260284, 0.997292637825012,
                0.996305286884308, 0.995250642299652, 0.994234919548035,
                0.993183135986328, 0.992080569267273, 0.990928232669830,
                0.989747583866119, 0.988448619842529, 0.987112879753113,
                0.985681176185608, 0.984169244766235, 0.982512056827545)
  Iethrisk <- c(0, 0,
                 0.25740993498319259, 0.61297954305717794, 0.33621598416696213,
                 0.15125173032243364, -0.17941562596577681, -0.35034236100577454,
                 -0.27783724832332168, -0.15927341226653660)
  Ismoke <- c(0, 0.21193771087603852, 0.66186343796859415,
               0.75707145871323056, 0.94962982514570360)

  dage  <- age / 10; dbmi <- bmi / 10
  age_1 <- sqrt(dage);      age_2 <- dage
  bmi_1 <- dbmi^(-2);       bmi_2 <- dbmi^(-2) * log(dbmi)

  age_1 <- age_1 - 2.086397409439087;  age_2 <- age_2 - 4.353054523468018
  bmi_1 <- bmi_1 - 0.152244374155998;  bmi_2 <- bmi_2 - 0.143282383680344
  rati  <- rati  - 3.506655454635620;  sbp   <- sbp   - 125.040039062500000
  town  <- town  - 0.416743695735931

  a <- Iethrisk[clamp_vec(ethrisk, 1L, 9L) + 1L] + Ismoke[clamp_vec(smoke_cat, 0L, 4L) + 1L]

  a <- a + age_1 * 4.4417863976316578   + age_2 * 0.028163721067299918
  a <- a + bmi_1 * 0.89423653047106633  + bmi_2 * (-6.5748047596104335)
  a <- a + rati  * 0.14339005616214209  + sbp   * 0.012897179584361372
  a <- a + town  * 0.066477263001143885
  a <- a + b_AF  * 1.6284780236484424   + b_ra  * 0.29012331040887707
  a <- a + b_renal * 1.0043796680368302 + b_treatedhyp * 0.61804305627881295
  a <- a + b_type1 * 1.8400348250874599 + b_type2 * 1.1711626412196512
  a <- a + fh_cvd  * 0.51472612036651955

  a <- a + age_1 * (smoke_cat==1) *  0.74644061443916665
  a <- a + age_1 * (smoke_cat==2) *  0.25685417118796666
  a <- a + age_1 * (smoke_cat==3) * -1.5452226707866523
  a <- a + age_1 * (smoke_cat==4) * -1.7113013709043405
  a <- a + age_1 * b_AF           * -7.0177986441269269
  a <- a + age_1 * b_renal        * -2.9684019256454390
  a <- a + age_1 * b_treatedhyp   * -4.2219906452967848
  a <- a + age_1 * b_type1        *  1.6835769546040080
  a <- a + age_1 * b_type2        * -2.9371798540034648
  a <- a + age_1 * bmi_1          *  0.17971962070446823
  a <- a + age_1 * bmi_2          * 40.242816676065814
  a <- a + age_1 * fh_cvd         *  0.14399792407539067
  a <- a + age_1 * sbp            * -0.036257523389977446
  a <- a + age_1 * town           *  0.37351380314334426

  a <- a + age_2 * (smoke_cat==1) * -0.19270577417482310
  a <- a + age_2 * (smoke_cat==2) * -0.15269650634589327
  a <- a + age_2 * (smoke_cat==3) *  0.23135639765214294
  a <- a + age_2 * (smoke_cat==4) *  0.23071650138682967
  a <- a + age_2 * b_AF           *  1.1395776028337732
  a <- a + age_2 * b_renal        *  0.43569632083309406
  a <- a + age_2 * b_treatedhyp   *  0.72659471088872396
  a <- a + age_2 * b_type1        * -0.63209777662756539
  a <- a + age_2 * b_type2        *  0.40232704348710868
  a <- a + age_2 * bmi_1          *  0.13192766227118777
  a <- a + age_2 * bmi_2          * -7.3211322435546409
  a <- a + age_2 * fh_cvd         * -0.13302600182737204
  a <- a + age_2 * sbp            *  0.0045842850495397955
  a <- a + age_2 * town           * -0.095237030084599078

  surv <- as.integer(clamp_vec(surv, 1L, 15L))
  100.0 * (1.0 - survivor[surv + 1L]^exp(a))
}

# ── QRISK2-2015 male ──────────────────────────────────────────────────────────
qrisk2_2015_male <- function(age, bmi, sbp, rati, town, smoke_cat, ethrisk,
                              b_AF=0, b_ra=0, b_renal=0, b_treatedhyp=0,
                              b_type1=0, b_type2=0, fh_cvd=0, surv=10L) {
  survivor <- c(0, 0.998205721378326, 0.996380507946014, 0.994511783123016,
                0.992520809173584, 0.990338504314423, 0.988223373889923,
                0.986053407192230, 0.983751118183136, 0.981297850608826,
                0.978794217109680, 0.976175725460052, 0.973336458206177,
                0.970354199409485, 0.967230498790741, 0.963918387889862)
  Iethrisk <- c(0, 0,
                 0.31733214304819191, 0.47385907860811155, 0.51713146559681455,
                 0.13703011573664192, -0.38855223049726639, -0.38124954853121945,
                 -0.40644613816509945, -0.22857155213773361)
  Ismoke <- c(0, 0.26844791581580202, 0.63076749738775917,
               0.71780788833786957, 0.87041725334654851)

  dage  <- age / 10; dbmi <- bmi / 10
  age_1 <- dage^(-1);       age_2 <- dage^2
  bmi_1 <- dbmi^(-2);       bmi_2 <- dbmi^(-2) * log(dbmi)

  age_1 <- age_1 - 0.233734160661697;  age_2 <- age_2 - 18.304403305053711
  bmi_1 <- bmi_1 - 0.146269768476486;  bmi_2 <- bmi_2 - 0.140587374567986
  rati  <- rati  - 4.321151256561279;  sbp   <- sbp   - 130.589752197265620
  town  <- town  - 0.551009356975555

  a <- Iethrisk[clamp_vec(ethrisk, 1L, 9L) + 1L] + Ismoke[clamp_vec(smoke_cat, 0L, 4L) + 1L]

  a <- a + age_1 * (-18.043731255037727) + age_2 * 0.023648645425430694
  a <- a + bmi_1 *   2.5388084343581578  + bmi_2 * (-9.1034725871528597)
  a <- a + rati  *   0.16843976361369095 + sbp   * 0.010500308938075482
  a <- a + town  *   0.032380163763448759
  a <- a + b_AF  *   1.0363048000259454  + b_ra  * 0.25199531347910126
  a <- a + b_renal * 0.83593528869952860 + b_treatedhyp * 0.66034596959178626
  a <- a + b_type1 * 1.3309170433446138  + b_type2 * 0.94543488927744179
  a <- a + fh_cvd  * 0.59860378971362815

  a <- a + age_1 * (smoke_cat==1) *  0.61868646993796839
  a <- a + age_1 * (smoke_cat==2) *  1.5522017055600055
  a <- a + age_1 * (smoke_cat==3) *  2.4407210657517648
  a <- a + age_1 * (smoke_cat==4) *  3.5140494491884624
  a <- a + age_1 * b_AF           *  8.0382925558108482
  a <- a + age_1 * b_renal        * -1.6389521229064483
  a <- a + age_1 * b_treatedhyp   *  8.4621771382346651
  a <- a + age_1 * b_type1        *  5.4977016563835504
  a <- a + age_1 * b_type2        *  3.3974747488766690
  a <- a + age_1 * bmi_1          * 33.848988101276760
  a <- a + age_1 * bmi_2          * (-140.67070254048971)
  a <- a + age_1 * fh_cvd         *  2.0858333154353321
  a <- a + age_1 * sbp            *  0.050128366883072054
  a <- a + age_1 * town           * -0.19882682171868507

  a <- a + age_2 * (smoke_cat==1) * -0.0040893975066796338
  a <- a + age_2 * (smoke_cat==2) * -0.0056065852346001768
  a <- a + age_2 * (smoke_cat==3) * -0.0018261006189440492
  a <- a + age_2 * (smoke_cat==4) * -0.0014997157296173290
  a <- a + age_2 * b_AF           *  0.0052471594895864343
  a <- a + age_2 * b_renal        * -0.017966358619354639
  a <- a + age_2 * b_treatedhyp   *  0.0092088445323379176
  a <- a + age_2 * b_type1        *  0.0047493510223424558
  a <- a + age_2 * b_type2        * -0.0048113775783491563
  a <- a + age_2 * bmi_1          *  0.062741075751394565
  a <- a + age_2 * bmi_2          * -0.23829149093857321
  a <- a + age_2 * fh_cvd         * -0.0049971149213281010
  a <- a + age_2 * sbp            * -0.000052370098795143509
  a <- a + age_2 * town           * -0.0012518116569283104

  surv <- as.integer(clamp_vec(surv, 1L, 15L))
  100.0 * (1.0 - survivor[surv + 1L]^exp(a))
}

qrisk2_2015 <- function(sex, age, bmi, sbp, rati, town,
                         smoke_cat, ethrisk,
                         b_AF=0, b_ra=0, b_renal=0,
                         b_treatedhyp=0, b_type1=0,
                         b_type2=0, fh_cvd=0, surv=10L) {
  sex1 <- toupper(substr(as.character(sex), 1, 1))
  out  <- rep(NA_real_, length(sex1))
  f    <- which(sex1 == "F")
  m    <- which(sex1 == "M")
  if (length(f))
    out[f] <- qrisk2_2015_female(age[f], bmi[f], sbp[f], rati[f], town[f],
                                  smoke_cat[f], ethrisk[f], b_AF[f], b_ra[f],
                                  b_renal[f], b_treatedhyp[f], b_type1[f],
                                  b_type2[f], fh_cvd[f], surv)
  if (length(m))
    out[m] <- qrisk2_2015_male(age[m], bmi[m], sbp[m], rati[m], town[m],
                                smoke_cat[m], ethrisk[m], b_AF[m], b_ra[m],
                                b_renal[m], b_treatedhyp[m], b_type1[m],
                                b_type2[m], fh_cvd[m], surv)
  out
}
```

```{r analysis_functions, include=FALSE}
# ── QRISK2 preparation (DM cohort) ────────────────────────────────────────────
prepare_qrisk_dm <- function(dm_df) {
  if (!requireNamespace("mice", quietly = TRUE))
    stop("prepare_qrisk_dm: 'mice' package required")

  town_center <- ifelse(dm_df$sex == "F", 0.416743695735931, 0.551009356975555)
  dm_df$townsend[is.na(dm_df$townsend)] <- town_center[is.na(dm_df$townsend)]

  imp_df <- data.frame(
    smk      = factor(dm_df$smoking_status,
                      levels = c("Non-smoker","Ex-smoker","Light","Moderate","Heavy"),
                      ordered = TRUE),
    bmi      = as.numeric(dm_df$bmi),
    sbp      = as.numeric(dm_df$sbp),
    tc_hdl   = as.numeric(dm_df$tc_hdl),
    age      = as.integer(dm_df$age),
    sex      = dm_df$sex,
    ldl      = as.numeric(dm_df$ldl),
    aht      = as.integer(dm_df$aht_flag),
    ckd      = as.integer(dm_df$ckd)
  )

  method          <- mice::make.method(imp_df)
  method["smk"]   <- "polr"
  method["bmi"]   <- "pmm"
  method["sbp"]   <- "pmm"
  method["tc_hdl"]<- "pmm"
  method[c("age","sex","ldl","aht","ckd")] <- ""

  n_miss <- c(smk=sum(is.na(imp_df$smk)), bmi=sum(is.na(imp_df$bmi)),
              sbp=sum(is.na(imp_df$sbp)), tc_hdl=sum(is.na(imp_df$tc_hdl)))
  if (any(n_miss > 0)) {
    set.seed(500L)
    imputed <- mice::complete(
      mice::mice(imp_df, m=1L, method=method, seed=500L, printFlag=FALSE)
    )
    dm_df$smoking_status <- as.character(imputed$smk)
    dm_df$bmi    <- imputed$bmi
    dm_df$sbp    <- imputed$sbp
    dm_df$tc_hdl <- imputed$tc_hdl
  }
  dm_df
}

# ── REACT score computation ────────────────────────────────────────────────────
compute_react_scores <- function(ascvd_df, model_file, surv_years=5L) {
  for (pkg in c("qs","msm","mice")) {
    if (!requireNamespace(pkg, quietly=TRUE))
      stop("compute_react_scores: '", pkg, "' package required")
  }
  message("[REACT] Loading MSM model: ", model_file)
  msm.cvd <- qs::qread(model_file)

  n_sex_na <- sum(is.na(ascvd_df$sex))
  if (n_sex_na > 0) {
    modal_sex <- names(sort(table(ascvd_df$sex), decreasing=TRUE))[1]
    ascvd_df$sex[is.na(ascvd_df$sex)] <- modal_sex
  }

  smk_react <- dplyr::case_when(
    ascvd_df$smoking_status == "Non-smoker" ~ "Non-smoker",
    ascvd_df$smoking_status == "Ex-smoker"  ~ "Ex-smoker",
    ascvd_df$smoking_status %in% c("Light","Moderate","Heavy") ~ "Active smoker",
    TRUE ~ NA_character_
  )

  imp_df <- data.frame(
    alf      = ascvd_df$alf,
    he       = as.integer(ascvd_df$hard_event),
    age      = as.integer(ascvd_df$age),
    sex      = ascvd_df$sex,
    type     = factor(case_when(ascvd_df$vascular_territory == 'Stroke' ~ 'ST',
                                TRUE ~ ascvd_df$vascular_territory),
                      levels=c("IHD","Stroke","PAD"), ordered=FALSE),
    wimd     = as.integer(ascvd_df$wimd),
    smk      = factor(smk_react,
                      levels=c("Non-smoker","Ex-smoker","Active smoker"), ordered=TRUE),
    dm       = as.integer(ascvd_df$dm_comorbid),
    dmta     = as.integer(ascvd_df$dmta),
    lvr      = as.integer(ascvd_df$liver_disease),
    htn      = as.integer(ascvd_df$htn),
    hf       = as.integer(ascvd_df$hf),
    nhdl_val = as.numeric(ascvd_df$non_hdl),
    egfr_val = as.numeric(ascvd_df$egfr),
    stringsAsFactors=FALSE
  )

  impute_vars        <- imp_df[, setdiff(names(imp_df), c("alf","he"))]
  method             <- mice::make.method(impute_vars)
  method["smk"]      <- "polr"
  method["nhdl_val"] <- "pmm"
  method["egfr_val"] <- "pmm"

  set.seed(500L)
  imputed <- mice::complete(
    mice::mice(impute_vars, m=1L, method=method, seed=500L, printFlag=FALSE)
  )
  imp_df[, names(imputed)] <- imputed

  imp_df$smk_msm <- factor(as.character(imp_df$smk),
                            levels=c("Non-smoker","Ex-smoker","Active smoker","Unknown"),
                            ordered=FALSE)
  imp_df$nhdl <- factor(dplyr::case_when(
    imp_df$nhdl_val <= 4.3 ~ "<= 4.3",
    imp_df$nhdl_val >  4.3 ~ "> 4.3"
  ), levels=c("<= 4.3","> 4.3"))
  imp_df$egfr <- factor(dplyr::case_when(
    imp_df$egfr_val >  60                         ~ "> 60",
    imp_df$egfr_val >= 30 & imp_df$egfr_val <= 60 ~ "30 - 60",
    imp_df$egfr_val <  30                         ~ "< 30"
  ), levels=c("> 60","30 - 60","< 30"))
  imp_df$age_cat <- factor(dplyr::case_when(
    imp_df$age <  50                        ~ "< 50",
    imp_df$age >= 50 & imp_df$age <= 59     ~ "50 - 59",
    imp_df$age >= 60 & imp_df$age <= 69     ~ "60 - 69",
    imp_df$age >= 70 & imp_df$age <= 79     ~ "70 - 79",
    imp_df$age >= 80                        ~ "80 - 85"
  ), levels=c("< 50","50 - 59","60 - 69","70 - 79","80 - 85"))

  n     <- nrow(imp_df)
  entry <- ifelse(imp_df$he == 1L, 3L, 1L)
  tick  <- max(1L, n %/% 20L)
  message(sprintf("[REACT] Computing passage probabilities for %d patients...", n))

  react_5y <- vapply(seq_len(n), function(i) {
    if (i %% tick == 0L)
      message(sprintf("[REACT]   %d / %d  (%.0f%%)", i, n, 100*i/n))
    passage <- msm::ppass.msm(
      msm.cvd, tot=surv_years,
      covariates=list(
        Age                      = imp_df$age_cat[i],
        Sex                      = imp_df$sex[i],
        Index_vascular_territory = imp_df$type[i],
        WIMD_Quintile            = imp_df$wimd[i],
        Smoker_status            = imp_df$smk_msm[i],
        Diabetes                 = imp_df$dm[i],
        Dementia                 = imp_df$dmta[i],
        Liver_disease            = imp_df$lvr[i],
        Hypertension             = imp_df$htn[i],
        Heart_failure            = imp_df$hf[i],
        non_HDL                  = imp_df$nhdl[i],
        eGFR                     = imp_df$egfr[i]
      )
    ) |> unclass() |> as.data.frame() |> tibble::rownames_to_column("State")
    prob <- if (entry[i] == 3L) passage$`State 4`[3L] else passage$`State 2`[1L]
    as.numeric(prob) * 100
  }, numeric(1L))

  data.frame(alf=imp_df$alf, react_5y_risk=react_5y)
}

# ── Cohort loader ──────────────────────────────────────────────────────────────
load_ehr_cohorts <- function(react_model_file=NULL) {
  if (!requireNamespace("sailr", quietly=TRUE))
    stop("sailr package required.")
  conn <- sailr::db2_open()
  on.exit(try(sailr::db2_close(conn), silent=TRUE))

  raw <- sailr::db2_run(conn, "
    SELECT
      ALF_PE      AS alf,
      COHORT      AS cohort,
      HE          AS hard_event,
      SEX         AS sex,
      AGE         AS age,
      ETHN        AS ethnicity_raw,
      SMK         AS smoking_raw,
      TYPE        AS diabetes_type,
      WIMD        AS wimd,
      TOWNSEND    AS townsend,
      BMI         AS bmi,
      SBP_VAL     AS sbp,
      LDL_VAL     AS ldl,
      TC_HDL      AS tc_hdl_ratio,
      NHDL_VAL    AS non_hdl,
      EGFR_VAL    AS egfr,
      DM          AS dm_comorbid,
      DMTA        AS dmta,
      LVR         AS liver_disease,
      HTN         AS htn,
      HF          AS hf,
      AF          AS af,
      RA          AS rheumatoid_arthritis,
      FH_CVD      AS fh_chd_u60,
      CKD         AS ckd,
      STATIN      AS statin_rx,
      LLT_INT     AS llt_int,
      AHT         AS aht_flag,
      AHT_INT     AS aht_int
    FROM SAILW1483V.CAPSTONE_RISK_CALCULATION_COHORT
  ")

  clean_cohort <- function(df) {
    df |>
      dplyr::mutate(
        ethnicity          = map_ehr_ethnicity(as.character(ethnicity_raw)),
        smoking_status     = map_ehr_smoking(as.character(smoking_raw)),
        sbp      = as.numeric(sbp),
        ldl      = as.numeric(ldl),
        tc_hdl   = as.numeric(tc_hdl_ratio),
        non_hdl  = as.numeric(non_hdl),
        bmi      = as.numeric(bmi),
        egfr     = as.numeric(egfr),
        wimd     = as.integer(wimd),
        townsend = as.numeric(townsend),
        statin_rx_last6m = as.integer(statin_rx),
        llt_int          = as.integer(llt_int),
        aht_int          = as.integer(aht_int),
        aht_flag         = as.integer(aht_flag),
        treated_htn      = as.integer(aht_flag),
        dm_type1 = as.integer(cohort == "DM" & !is.na(diabetes_type) &
                                toupper(diabetes_type) == "T1"),
        dm_type2 = as.integer(cohort == "DM" & (is.na(diabetes_type) |
                                toupper(diabetes_type) %in% c("T2","U"))),
        vascular_territory = dplyr::case_when(
          cohort == "ASCVD" & toupper(diabetes_type) == "IHD"    ~ "IHD",
          cohort == "ASCVD" & toupper(diabetes_type) == "STROKE" ~ "Stroke",
          cohort == "ASCVD" & toupper(diabetes_type) == "PAD"    ~ "PAD",
          TRUE ~ NA_character_
        ),
        ckd                  = as.integer(ckd),
        hf                   = as.integer(hf),
        af                   = as.integer(af),
        rheumatoid_arthritis = as.integer(rheumatoid_arthritis),
        fh_chd_u60           = as.integer(fh_chd_u60),
        liver_disease        = as.integer(liver_disease),
        htn                  = as.integer(htn),
        dm_comorbid          = as.integer(dm_comorbid),
        dmta                 = as.integer(dmta),
        hard_event           = as.integer(hard_event),
        age                  = as.integer(age)
      ) |>
      dplyr::select(-ethnicity_raw, -smoking_raw, -tc_hdl_ratio)
  }

  all_data <- clean_cohort(raw)
  dm    <- dplyr::filter(all_data, cohort == "DM")
  ascvd <- dplyr::filter(all_data, cohort == "ASCVD")

  dm <- prepare_qrisk_dm(dm)

  react_scores        <- compute_react_scores(ascvd, model_file=react_model_file)
  ascvd$react_5y_risk <- react_scores$react_5y_risk[match(ascvd$alf, react_scores$alf)]

  message(sprintf("[ehr] DM cohort: %d patients | ASCVD cohort: %d patients",
                  nrow(dm), nrow(ascvd)))
  list(dm=dm, ascvd=ascvd)
}

# ── QRISK2 recalculation for DM agents ────────────────────────────────────────
recalc_qrisk_dm <- function(df) {
  dm <- df$cohort == "DM"
  if (!any(dm)) return(df)
  sub <- df[dm, ]
  smoke_cat <- map_smoke_cat(sub$smoking_status)
  smoke_cat[is.na(smoke_cat)] <- 0L
  rati <- pmax(as.numeric(sub$tc_hdl), 1.0)
  b_treatedhyp <- as.integer(sub$aht_int >= 1)
  town <- as.numeric(sub$townsend); town[is.na(town)] <- 0.0

  df$qrisk10y[dm] <- qrisk2_2015(
    sex=sub$sex, age=as.numeric(sub$age), bmi=sub$bmi, sbp=sub$sbp,
    rati=rati, town=town, smoke_cat=smoke_cat,
    ethrisk=map_ethrisk(sub$ethnicity),
    b_AF=sub$af, b_ra=sub$rheumatoid_arthritis, b_renal=sub$ckd,
    b_treatedhyp=b_treatedhyp,
    b_type1=sub$dm_type1, b_type2=sub$dm_type2,
    fh_cvd=sub$fh_chd_u60
  )
  df
}

# ── Compute baseline risk scores ───────────────────────────────────────────────
# Adds qrisk10y and react5y columns to a combined cohort data frame.
compute_risk_scores <- function(dm_df, ascvd_df) {
  dm_df$cohort    <- "DM"
  ascvd_df$cohort <- "ASCVD"
  df <- dplyr::bind_rows(dm_df, ascvd_df)

  df$qrisk10y <- NA_real_
  df$react5y  <- ifelse(df$cohort == "ASCVD", df$react_5y_risk, NA_real_)
  df <- recalc_qrisk_dm(df)

  list(
    dm    = df[df$cohort == "DM", ],
    ascvd = df[df$cohort == "ASCVD", ]
  )
}

# ── Baseline characteristics table ────────────────────────────────────────────
make_baseline_table <- function(dm_df, ascvd_df) {
  overall <- dplyr::bind_rows(dm_df, ascvd_df)

  pct <- function(x)   sprintf("%d (%.1f%%)", sum(x, na.rm=TRUE), 100*mean(x, na.rm=TRUE))
  mn  <- function(x)   sprintf("%.1f (%.1f)", mean(x, na.rm=TRUE), sd(x, na.rm=TRUE))
  mn2 <- function(x)   sprintf("%.2f (%.2f)", mean(x, na.rm=TRUE), sd(x, na.rm=TRUE))

  summ <- function(df, cohort="overall") {
    is_dm <- cohort == "DM"
    vt <- function(val) {
      if (is_dm) return("-")
      pct(df$vascular_territory == val)
    }
    list(
      "N"                               = nrow(df),
      "Age (years)"                     = mn(df$age),
      "Male sex"                        = pct(df$sex == "M"),
      "Systolic blood pressure (mmHg)"  = mn(df$sbp),
      "Body mass index (kg/m²)"         = mn(df$bmi),
      "LDL cholesterol (mmol/L)"        = mn2(df$ldl),
      "Statin in last 6 months"         = pct(df$statin_rx_last6m),
      "Treated hypertension"            = pct(df$treated_htn),
      "Atrial fibrillation"             = pct(df$af),
      "CKD stage 3-5"                   = pct(df$ckd),
      "Rheumatoid arthritis"            = pct(df$rheumatoid_arthritis),
      "Family history of CHD <60 years" = pct(df$fh_chd_u60),
      "Townsend deprivation score"      = mn(df$townsend),
      "IHD"                             = vt("IHD"),
      "Stroke"                          = vt("Stroke"),
      "Peripheral arterial disease"     = vt("PAD"),
      "Diabetes mellitus"               = if (is_dm) "-" else pct(df$cohort == "DM" | df$dm_comorbid == 1L),
      "Heart failure"                   = pct(df$hf),
      "Liver disease"                   = pct(df$liver_disease),
      "Dementia"                        = pct(df$dmta),
      "Non-HDL-C ≤4.3 mmol/L"          = pct(df$non_hdl <= 4.3),
      "eGFR >60 mL/min/1.73m²"         = pct(df$egfr > 60),
      "WIMD quintile 1 (most deprived)" = pct(df$wimd == 1L)
    )
  }

  so <- summ(overall, "overall")
  sa <- summ(ascvd_df, "ASCVD")
  sd_ <- summ(dm_df,   "DM")

  data.frame(
    Characteristic = names(so),
    Overall        = unlist(so),
    ASCVD          = unlist(sa),
    DM             = unlist(sd_),
    stringsAsFactors=FALSE, row.names=NULL
  )
}

# ── Treatment rate table ───────────────────────────────────────────────────────
# Proportion of patients at each LLT intensity (0–4) and AHT intensity (0–4),
# by cohort. Returns a long data frame suitable for stacked bar charts.
compute_treatment_rates <- function(dm_df, ascvd_df) {
  cohorts <- list(
    list(name="DM (primary)",      df=dm_df),
    list(name="ASCVD (secondary)", df=ascvd_df)
  )
  llt_labs <- c("0"="No treatment","1"="Other treatment","2"="LI statin",
                "3"="HI statin","4"="Combination therapy")
  aht_labs <- c("0"="No treatment","1"="1 drug class","2"="2 classes",
                "3"="3 classes","4"="4 classes")

  dplyr::bind_rows(lapply(cohorts, function(co) {
    df <- co$df
    llt <- table(factor(df$llt_int, levels=0:4)) / nrow(df) * 100
    aht <- table(factor(df$aht_int, levels=0:4)) / nrow(df) * 100
    dplyr::bind_rows(
      data.frame(Cohort=co$name, Treatment="LLT",
                 Intensity=names(llt_labs), Label=llt_labs,
                 Pct=as.numeric(llt), stringsAsFactors=FALSE),
      data.frame(Cohort=co$name, Treatment="AHT",
                 Intensity=names(aht_labs), Label=aht_labs,
                 Pct=as.numeric(aht), stringsAsFactors=FALSE)
    )
  }))
}

# ── Treatment gap ──────────────────────────────────────────────────────────────
compute_treatment_gap <- function(dm_df, ascvd_df) {
  ldl_target_dm    <- 2.0
  ldl_target_ascvd <- 1.8
  sbp_target       <- 130.0
  specs <- list(
    list(cname="DM",    df=dm_df,    ldl_tgt=ldl_target_dm),
    list(cname="ASCVD", df=ascvd_df, ldl_tgt=ldl_target_ascvd)
  )
  dplyr::bind_rows(lapply(specs, function(sp) {
    df <- sp$df
    data.frame(
      Cohort                    = sp$cname,
      N                         = nrow(df),
      `LDL at target (%)`       = round(100*mean(df$ldl <= sp$ldl_tgt, na.rm=TRUE), 1),
      `Mean LDL gap (mmol/L)`   = round(mean(pmax(df$ldl - sp$ldl_tgt, 0), na.rm=TRUE), 2),
      `On LLT (%)`           = round(100*mean(df$statin_rx_last6m, na.rm=TRUE), 1),
      `SBP at target (%)`       = round(100*mean(df$sbp <= sbp_target, na.rm=TRUE), 1),
      `Mean SBP gap (mmHg)`     = round(mean(pmax(df$sbp - sbp_target, 0), na.rm=TRUE), 1),
      `On AHT (%)` = round(100*mean(df$treated_htn, na.rm=TRUE), 1),
      check.names=FALSE, stringsAsFactors=FALSE
    )
  }))
}

# ── Detailed treatment gap breakdown ──────────────────────────────────────────
compute_treatment_gap_detailed <- function(dm_df, ascvd_df) {
  ldl_target_dm <- 2.0; ldl_target_ascvd <- 1.8; sbp_target <- 130.0
  specs <- list(
    list(cname="DM",    df=dm_df,    ldl_tgt=ldl_target_dm),
    list(cname="ASCVD", df=ascvd_df, ldl_tgt=ldl_target_ascvd)
  )
  ldl_rows <- dplyr::bind_rows(lapply(specs, function(sp) {
    df  <- sp$df; tgt <- sp$ldl_tgt
    on_llt <- df$llt_int > 0L
    at_tgt <- !is.na(df$ldl) & df$ldl <= tgt
    pct <- function(x) round(100 * mean(x, na.rm=TRUE), 1)
    data.frame(
      Cohort   = sp$cname,
      Category = c("No LLT, at target", "LLT, at target",
                   "LLT, not at target", "No LLT, not at target"),
      N   = c(sum(!on_llt &  at_tgt, na.rm=TRUE), sum( on_llt &  at_tgt, na.rm=TRUE),
              sum( on_llt & !at_tgt, na.rm=TRUE), sum(!on_llt & !at_tgt, na.rm=TRUE)),
      Pct = c(pct(!on_llt &  at_tgt), pct( on_llt &  at_tgt),
              pct( on_llt & !at_tgt), pct(!on_llt & !at_tgt)),
      stringsAsFactors = FALSE
    )
  }))
  sbp_rows <- dplyr::bind_rows(lapply(specs, function(sp) {
    df  <- sp$df
    at_tgt <- !is.na(df$sbp) & df$sbp <= sbp_target
    pct <- function(x) round(100 * mean(x, na.rm=TRUE), 1)
    data.frame(
      Cohort   = sp$cname,
      Category = c("At target", "1-3 agents, above target",
                   "4 agents, above target", "No AHT, above target"),
      N   = c(sum( at_tgt, na.rm=TRUE),
              sum(!at_tgt & df$aht_int %in% 1:3, na.rm=TRUE),
              sum(!at_tgt & df$aht_int == 4L,    na.rm=TRUE),
              sum(!at_tgt & df$aht_int == 0L,    na.rm=TRUE)),
      Pct = c(pct( at_tgt), pct(!at_tgt & df$aht_int %in% 1:3),
              pct(!at_tgt & df$aht_int == 4L), pct(!at_tgt & df$aht_int == 0L)),
      stringsAsFactors = FALSE
    )
  }))
  list(ldl=ldl_rows, sbp=sbp_rows)
}

# ── Target attainment and risk projection table ────────────────────────────────
# Uses meta-analytic RRs to project how risk changes under each scenario without
# running the ABM. Based on HR = RR_SBP^(ΔSBP/10) × RR_LDL^ΔLDL.
#
# Key design: project_risk() takes simultaneous deltas for BOTH cohorts and
# returns one qrisk and one react value. The outer loop builds per-cohort biomarker
# rows (Mean, Delta, % at target) but shares the same risk projection across
# both cohort rows for a given scenario - avoiding the bug where dplyr::first()
# in reshape_tbl would pick up whichever cohort happened to be listed first.
make_target_tables <- function(dm_df, ascvd_df) {
  rr_sbp_dm    <- 0.90; rr_ldl_dm    <- 0.79
  rr_sbp_ascvd <- 0.80; rr_ldl_ascvd <- 0.78
  ldl_target_dm <- 2.0; ldl_target_ascvd <- 1.8; sbp_target <- 130.0

  scored  <- compute_risk_scores(dm_df, ascvd_df)
  dm_s    <- scored$dm; ascvd_s <- scored$ascvd

  project_risk <- function(d_sbp_dm=0, d_ldl_dm=0, d_sbp_ascvd=0, d_ldl_ascvd=0) {
    .pr <- function(s_dm, l_dm, s_asc, l_asc) {
      hr_dm    <- s_dm^(d_sbp_dm/10)     * l_dm^d_ldl_dm
      hr_ascvd <- s_asc^(d_sbp_ascvd/10) * l_asc^d_ldl_ascvd
      list(
        qrisk = round(mean(clamp_vec(dm_s$qrisk10y   * hr_dm,    0, 100), na.rm=TRUE), 1),
        react = round(mean(clamp_vec(ascvd_s$react5y * hr_ascvd, 0, 100), na.rm=TRUE), 1)
      )
    }
    cen <- .pr(0.90, 0.79, 0.80, 0.78)
    pes <- .pr(0.95, 0.86, 0.83, 0.80)
    opt <- .pr(0.84, 0.72, 0.77, 0.76)
    list(qrisk=cen$qrisk, react=cen$react,
         qrisk_lo=opt$qrisk, qrisk_hi=pes$qrisk,
         react_lo=opt$react, react_hi=pes$react)
  }

  ldl_dm    <- dm_df$ldl;    ldl_ascvd <- ascvd_df$ldl
  sbp_dm    <- dm_df$sbp;    sbp_ascvd <- ascvd_df$sbp

  rows_ldl <- list()
  for (cov in c(0, 0.10, 0.25, 0.50, 0.75, 0.90, 1.00)) {
    d_dm    <- cov * mean(pmax(ldl_dm    - ldl_target_dm,    0), na.rm=TRUE)
    d_ascvd <- cov * mean(pmax(ldl_ascvd - ldl_target_ascvd, 0), na.rm=TRUE)
    rs      <- project_risk(d_ldl_dm=d_dm, d_ldl_ascvd=d_ascvd)
    pct_dm    <- 100 * (mean(ldl_dm    <= ldl_target_dm,    na.rm=TRUE) +
                        cov * mean(ldl_dm    > ldl_target_dm,    na.rm=TRUE))
    pct_ascvd <- 100 * (mean(ldl_ascvd <= ldl_target_ascvd, na.rm=TRUE) +
                        cov * mean(ldl_ascvd > ldl_target_ascvd, na.rm=TRUE))
    scen <- if (cov == 0) "Baseline" else sprintf("%d%% at target", as.integer(cov * 100))
    for (info in list(list("DM", ldl_dm, d_dm, pct_dm),
                      list("ASCVD", ldl_ascvd, d_ascvd, pct_ascvd))) {
      rows_ldl[[length(rows_ldl)+1]] <- data.frame(
        Cohort=info[[1]], Scenario=scen,
        Mean=round(mean(info[[2]], na.rm=TRUE) - info[[3]], 2), Delta=round(info[[3]], 2),
        pct_at_target=round(info[[4]], 1),
        mean_qrisk=rs$qrisk, mean_react=rs$react,
        qrisk_lo=rs$qrisk_lo, qrisk_hi=rs$qrisk_hi,
        react_lo=rs$react_lo, react_hi=rs$react_hi)
    }
  }

  rows_sbp <- list()
  for (cov in c(0, 0.10, 0.25, 0.50, 0.75, 0.90, 1.00)) {
    d_dm    <- cov * mean(pmax(sbp_dm    - sbp_target, 0), na.rm=TRUE)
    d_ascvd <- cov * mean(pmax(sbp_ascvd - sbp_target, 0), na.rm=TRUE)
    rs      <- project_risk(d_sbp_dm=d_dm, d_sbp_ascvd=d_ascvd)
    pct_dm    <- 100 * (mean(sbp_dm    <= sbp_target, na.rm=TRUE) +
                        cov * mean(sbp_dm    > sbp_target, na.rm=TRUE))
    pct_ascvd <- 100 * (mean(sbp_ascvd <= sbp_target, na.rm=TRUE) +
                        cov * mean(sbp_ascvd > sbp_target, na.rm=TRUE))
    scen <- if (cov == 0) "Baseline" else sprintf("%d%% at target", as.integer(cov * 100))
    for (info in list(list("DM", sbp_dm, d_dm, pct_dm),
                      list("ASCVD", sbp_ascvd, d_ascvd, pct_ascvd))) {
      rows_sbp[[length(rows_sbp)+1]] <- data.frame(
        Cohort=info[[1]], Scenario=scen,
        Mean=round(mean(info[[2]], na.rm=TRUE) - info[[3]], 1), Delta=round(info[[3]], 1),
        pct_at_target=round(info[[4]], 1),
        mean_qrisk=rs$qrisk, mean_react=rs$react,
        qrisk_lo=rs$qrisk_lo, qrisk_hi=rs$qrisk_hi,
        react_lo=rs$react_lo, react_hi=rs$react_hi)
    }
  }

  reshape_tbl <- function(rows, biomarker, decimals=2) {
    df <- do.call(rbind, rows)
    scen_order <- unique(df$Scenario)
    base_pct <- stats::setNames(
      df$pct_at_target[df$Scenario == "Baseline"],
      df$Cohort[df$Scenario == "Baseline"]
    )
    df$delta_pct <- df$pct_at_target - base_pct[df$Cohort]
    df$pct_str <- ifelse(df$Scenario == "Baseline",
      sprintf("%.1f%%", df$pct_at_target),
      sprintf("%.1f%%<br>(+%.1f%%)", df$pct_at_target, df$delta_pct))
    fmt     <- function(x, d=decimals) formatC(x, format="f", digits=d)
    fmt_neg <- function(x, d=decimals) ifelse(x == 0, fmt(0, d), paste0("-", fmt(x, d)))
    df$mean_str <- ifelse(df$Scenario == "Baseline", fmt(df$Mean),
      paste0(fmt(df$Mean), "<br>(", fmt_neg(df$Delta), ")"))
    risk_cols <- df |>
      dplyr::group_by(Scenario) |>
      dplyr::summarise(
        qrisk=dplyr::first(mean_qrisk), react=dplyr::first(mean_react),
        qrisk_lo=dplyr::first(qrisk_lo), qrisk_hi=dplyr::first(qrisk_hi),
        react_lo=dplyr::first(react_lo), react_hi=dplyr::first(react_hi),
        .groups="drop")
    qr_base <- risk_cols$qrisk[risk_cols$Scenario == "Baseline"]
    re_base <- risk_cols$react[risk_cols$Scenario == "Baseline"]
    fmt1 <- function(x) formatC(x, format="f", digits=1)
    is_base <- risk_cols$Scenario == "Baseline"
    risk_cols$`QRISK2 10y (%)` <- ifelse(is_base, fmt1(risk_cols$qrisk),
      paste0(fmt1(risk_cols$qrisk), "<br>(",
             ifelse(qr_base-risk_cols$qrisk==0, fmt1(0), paste0("-",fmt1(qr_base-risk_cols$qrisk))), ")"))
    risk_cols$`QRISK2 95% CI` <- ifelse(is_base, "—",
      paste0(fmt1(risk_cols$qrisk_lo), "–", fmt1(risk_cols$qrisk_hi)))
    risk_cols$`REACT 5y (%)` <- ifelse(is_base, fmt1(risk_cols$react),
      paste0(fmt1(risk_cols$react), "<br>(",
             ifelse(re_base-risk_cols$react==0, fmt1(0), paste0("-",fmt1(re_base-risk_cols$react))), ")"))
    risk_cols$`REACT 95% CI` <- ifelse(is_base, "—",
      paste0(fmt1(risk_cols$react_lo), "–", fmt1(risk_cols$react_hi)))
    risk_cols <- risk_cols[, c("Scenario","QRISK2 10y (%)","QRISK2 95% CI","REACT 5y (%)","REACT 95% CI")]
    wide <- tidyr::pivot_wider(
      df[, c("Cohort","Scenario","mean_str","pct_str")],
      id_cols="Scenario", names_from="Cohort",
      values_from=c("mean_str","pct_str"))
    wide <- wide[, c("Scenario","mean_str_DM","pct_str_DM","mean_str_ASCVD","pct_str_ASCVD")]
    names(wide) <- c("Scenario","DM Mean (Δ)","DM % at target","ASCVD Mean (Δ)","ASCVD % at target")
    wide <- dplyr::left_join(wide, risk_cols, by="Scenario")
    wide <- wide[match(scen_order[!duplicated(scen_order)], wide$Scenario), ]
    wide <- wide[, c("Scenario","DM Mean (Δ)","DM % at target","QRISK2 10y (%)","QRISK2 95% CI",
                     "ASCVD Mean (Δ)","ASCVD % at target","REACT 5y (%)","REACT 95% CI")]
    wide$biomarker <- biomarker
    wide
  }

  ldl_tbl <- reshape_tbl(rows_ldl, "LDL-C (mmol/L)", decimals=2)
  sbp_tbl <- reshape_tbl(rows_sbp, "SBP (mmHg)",     decimals=1)
  list(ldl=ldl_tbl, sbp=sbp_tbl, combined=dplyr::bind_rows(ldl_tbl, sbp_tbl))
}

# ── Events prevented analysis ──────────────────────────────────────────────────
# For each combination of risk quartile and intervention level, estimates the
# number of 5-year CVD events prevented per 1,000 patients using the meta-analytic
# RR formula: events_prevented = N × baseline_5yr_risk × (1 - HR)
# where HR = RR_SBP^(ΔSBP/10) × RR_LDL^ΔLDL.
#
# QRISK2 is a 10-year score; conversion to 5-year: p_5yr = 1 - (1-p_10yr)^0.5
compute_opportunity <- function(dm_df, ascvd_df, dm_scores, ascvd_scores) {
  rr_sbp_dm    <- 0.90; rr_ldl_dm    <- 0.79
  rr_sbp_ascvd <- 0.80; rr_ldl_ascvd <- 0.78

  # Intervention levels: (label, d_sbp_mmhg, d_ldl_mmol)
  interventions <- list(
    list(label="Mild\n(5mmHg / 0.3mmol)", d_sbp=5,  d_ldl=0.3),
    list(label="Moderate\n(10mmHg / 0.6mmol)", d_sbp=10, d_ldl=0.6),
    list(label="Intensive\n(15mmHg / 1.0mmol)", d_sbp=15, d_ldl=1.0)
  )

  # DM quartiles based on QRISK2 10yr, then converted to 5yr
  q_dm <- quantile(dm_scores$qrisk10y, probs=c(0,0.25,0.5,0.75,1), na.rm=TRUE)
  dm_merged <- dm_df
  dm_merged$qrisk10y   <- dm_scores$qrisk10y[match(dm_df$alf, dm_scores$alf)]
  dm_merged$risk_q     <- cut(dm_merged$qrisk10y, breaks=q_dm,
                               labels=c("Q1 (lowest)","Q2","Q3","Q4 (highest)"),
                               include.lowest=TRUE)
  dm_merged$risk5y     <- 1 - (1 - dm_merged$qrisk10y/100)^0.5

  q_ascvd <- quantile(ascvd_scores$react5y, probs=c(0,0.25,0.5,0.75,1), na.rm=TRUE)
  ascvd_merged <- ascvd_df
  ascvd_merged$react5y  <- ascvd_scores$react5y[match(ascvd_df$alf, ascvd_scores$alf)]
  ascvd_merged$risk_q   <- cut(ascvd_merged$react5y, breaks=q_ascvd,
                                 labels=c("Q1 (lowest)","Q2","Q3","Q4 (highest)"),
                                 include.lowest=TRUE)
  ascvd_merged$risk5y   <- ascvd_merged$react5y / 100

  dplyr::bind_rows(lapply(interventions, function(iv) {
    hr_dm    <- rr_sbp_dm^(iv$d_sbp/10)    * rr_ldl_dm^iv$d_ldl
    hr_ascvd <- rr_sbp_ascvd^(iv$d_sbp/10) * rr_ldl_ascvd^iv$d_ldl

    dm_res <- dm_merged |>
      dplyr::filter(!is.na(risk_q)) |>
      dplyr::group_by(risk_q) |>
      dplyr::summarise(
        events_prevented_per1k = mean(risk5y, na.rm=TRUE) * (1 - hr_dm) * 1000,
        n = dplyr::n(), .groups="drop"
      ) |>
      dplyr::mutate(Cohort="DM (primary)", Intervention=iv$label)

    ascvd_res <- ascvd_merged |>
      dplyr::filter(!is.na(risk_q)) |>
      dplyr::group_by(risk_q) |>
      dplyr::summarise(
        events_prevented_per1k = mean(risk5y, na.rm=TRUE) * (1 - hr_ascvd) * 1000,
        n = dplyr::n(), .groups="drop"
      ) |>
      dplyr::mutate(Cohort="ASCVD (secondary)", Intervention=iv$label)

    dplyr::bind_rows(dm_res, ascvd_res)
  }))
}

# ── 3D opportunity landscape matrix ───────────────────────────────────────────
# Returns a matrix of events prevented per 1,000 patients, indexed by
# risk quartile (rows) and SBP reduction band (columns).
# Used for the plotly surface plot - DM and ASCVD produced separately.
compute_sbp_risk_matrix <- function(scored_df, cohort_type) {
  rr_sbp    <- if (cohort_type == "DM") 0.90 else 0.80
  rr_ldl    <- if (cohort_type == "DM") 0.79 else 0.78
  risk_col  <- if (cohort_type == "DM") "qrisk10y" else "react5y"
  to_5yr    <- cohort_type == "DM"   # QRISK2 is 10-yr; REACT is already 5-yr

  sbp_reductions <- c(0, 2.5, 5, 7.5, 10, 12.5, 15, 20)
  ldl_reductions <- sbp_reductions * 0.06  # co-vary LDL proportionally (mild assumption)

  q <- quantile(scored_df[[risk_col]], probs=c(0,0.25,0.5,0.75,1), na.rm=TRUE)
  scored_df$risk_q <- cut(scored_df[[risk_col]], breaks=q,
                           labels=c("Q1","Q2","Q3","Q4"),
                           include.lowest=TRUE)

  risk5y_by_q <- scored_df |>
    dplyr::filter(!is.na(risk_q)) |>
    dplyr::group_by(risk_q) |>
    dplyr::summarise(
      mean_risk5y = {
        r <- mean(.data[[risk_col]], na.rm=TRUE) / 100
        if (to_5yr) 1 - (1 - r)^0.5 else r
      },
      .groups="drop"
    )

  mat <- outer(
    risk5y_by_q$mean_risk5y,
    seq_along(sbp_reductions),
    FUN=function(risk, idx) {
      d_sbp <- sbp_reductions[idx]
      d_ldl <- ldl_reductions[idx]
      hr    <- rr_sbp^(d_sbp/10) * rr_ldl^d_ldl
      risk * (1 - hr) * 1000
    }
  )
  rownames(mat) <- as.character(risk5y_by_q$risk_q)
  colnames(mat) <- paste0(sbp_reductions, "mmHg")
  mat
}
```

```{r theme, include=FALSE}
theme_cap <- function() {
  theme_minimal(base_size=11) +
    theme(legend.position="bottom",
          legend.title=element_text(face="bold"),
          panel.grid.minor=element_blank(),
          strip.text=element_text(face="bold"))
}

dx_pal <- c("Stable diagnosis"="#56026a", "MACE diagnosis"="#34aa6f")
```

```{r run_analysis, include=FALSE}
# Use a separate cache file per data source so switching doesn't require
# manual cache deletion.
.desc_cache_file <- file.path(
  cache_dir,
  if (data_source == "synthetic") "descriptive_synthetic.rds" else "descriptive.rds"
)

desc_cache <- cache_or_run(
  .desc_cache_file,
  function() {
    .prog <- function(...) message(sprintf("[%s] %s", format(Sys.time(), "%H:%M:%S"), paste0(...)))

    # ── Load cohorts ───────────────────────────────────────────────────────────
    if (data_source == "synthetic") {
      .prog("Step 1/6 - loading synthetic cohorts from ", synthetic_cohorts_file)
      .ext <- tools::file_ext(synthetic_cohorts_file)
      syn  <- if (.ext == "qs") qs::qread(synthetic_cohorts_file)
              else               readRDS(synthetic_cohorts_file)
      dm_base    <- syn$dm
      ascvd_base <- syn$ascvd
      .prog("  DM n=", nrow(dm_base), "  ASCVD n=", nrow(ascvd_base))
    } else {
      # Reuse main analysis cache to avoid re-running ~15,000 ppass.msm calls
      main_cache_file <- file.path("cache", "analysis.rds")
      if (file.exists(main_cache_file)) {
        .prog("Step 1/6 - loading cohorts from main analysis cache")
        main_cache <- readRDS(main_cache_file)
        dm_base    <- main_cache$dm_base
        ascvd_base <- main_cache$ascvd_base
        rm(main_cache)
      } else {
        .prog("Step 1/6 - loading cohorts from SAIL (main cache not found)")
        cohorts    <- load_ehr_cohorts(react_model_file=react_model_file)
        dm_base    <- cohorts$dm
        ascvd_base <- cohorts$ascvd
      }
    }
    
    .prog("Step 2/6 - baseline table")
    baseline_table <- make_baseline_table(dm_base, ascvd_base)

    .prog("Step 3/6 - risk scores")
    scored      <- compute_risk_scores(dm_base, ascvd_base)
    dm_scored   <- scored$dm
    ascvd_scored <- scored$ascvd

    .prog("Step 4/6 - treatment rates and gap")
    treatment_rates <- compute_treatment_rates(dm_base, ascvd_base)
    treatment_gap   <- compute_treatment_gap(dm_base, ascvd_base)

    .prog("Step 5/6 - biomarker target tables")
    target_tbls     <- make_target_tables(dm_base, ascvd_base)
    biomarker_table <- target_tbls$combined

    .prog("Step 6/6 - opportunity analysis")
    opportunity_df <- compute_opportunity(dm_base, ascvd_base, dm_scored, ascvd_scored)
    mat_dm    <- compute_sbp_risk_matrix(dm_scored,    "DM")
    mat_ascvd <- compute_sbp_risk_matrix(ascvd_scored, "ASCVD")

    list(
      dm_base=dm_base, ascvd_base=ascvd_base,
      dm_scored=dm_scored, ascvd_scored=ascvd_scored,
      baseline_table=baseline_table,
      treatment_rates=treatment_rates,
      treatment_gap=treatment_gap,
      biomarker_table=biomarker_table,
      opportunity_df=opportunity_df,
      mat_dm=mat_dm, mat_ascvd=mat_ascvd
    )
  }
)

list2env(desc_cache, envir=environment())
ldl_target_dm <- 2.0; ldl_target_ascvd <- 1.8; sbp_target <- 130.0
```

```{r, include = FALSE}

qsave(dm_base, 'dm_base.qs')
qsave(ascvd_base, 'ascvd_base.qs')  

```


```{r inline_values, include=FALSE}
# ── Inline scalar values for narrative text ────────────────────────────────────
# All placeholders in the results paragraphs are populated from this list (iv).
rr_sbp_dm <- 0.90; rr_ldl_dm <- 0.79
rr_sbp_ascvd <- 0.80; rr_ldl_ascvd <- 0.78

.hr <- function(d_sbp=0, d_ldl=0, cohort=c("DM","ASCVD")) {
  cohort <- match.arg(cohort)
  if (cohort=="DM")    rr_sbp_dm^(d_sbp/10)    * rr_ldl_dm^d_ldl
  else                 rr_sbp_ascvd^(d_sbp/10) * rr_ldl_ascvd^d_ldl
}
.mn <- function(x) mean(x, na.rm=TRUE)
.pct <- function(x, digits=1) round(100 * .mn(x), digits)

iv <- local({

  # ── Cohort sizes ──────────────────────────────────────────────────────────
  n_dm    <- nrow(dm_base);  n_ascvd <- nrow(ascvd_base)

  # ── DM demographics ───────────────────────────────────────────────────────
  mean_age_dm   <- round(.mn(dm_base$age), 1)
  pct_male_dm   <- .pct(dm_base$sex == "M")
  pct_htn_dm    <- .pct(dm_base$htn == 1L)
  pct_hf_dm     <- .pct(dm_base$hf  == 1L)
  pct_wimd1_dm  <- .pct(dm_base$wimd == 1L)
  pct_statin_dm <- .pct(dm_base$statin_rx_last6m == 1L)
  mean_ldl_dm   <- round(.mn(dm_base$ldl), 2)
  sd_ldl_dm     <- round(sd(dm_base$ldl, na.rm=TRUE), 2)
  mean_sbp_dm   <- round(.mn(dm_base$sbp), 1)
  sd_sbp_dm     <- round(sd(dm_base$sbp, na.rm=TRUE), 1)

  # ── ASCVD demographics ────────────────────────────────────────────────────
  mean_age_ascvd   <- round(.mn(ascvd_base$age), 1)
  pct_male_ascvd   <- .pct(ascvd_base$sex == "M")
  pct_ihd          <- .pct(ascvd_base$vascular_territory == "IHD")
  pct_stroke       <- .pct(ascvd_base$vascular_territory == "Stroke")
  pct_pad          <- .pct(ascvd_base$vascular_territory == "PAD")
  pct_htn_ascvd    <- .pct(ascvd_base$htn == 1L)
  pct_hf_ascvd     <- .pct(ascvd_base$hf  == 1L)
  pct_wimd1_ascvd  <- .pct(ascvd_base$wimd == 1L)
  pct_statin_ascvd <- .pct(ascvd_base$statin_rx_last6m == 1L)
  mean_ldl_ascvd   <- round(.mn(ascvd_base$ldl), 2)
  sd_ldl_ascvd     <- round(sd(ascvd_base$ldl, na.rm=TRUE), 2)
  mean_sbp_ascvd   <- round(.mn(ascvd_base$sbp), 1)
  sd_sbp_ascvd     <- round(sd(ascvd_base$sbp, na.rm=TRUE), 1)

  # ── Risk scores ───────────────────────────────────────────────────────────
  mean_qrisk     <- round(.mn(dm_scored$qrisk10y), 1)
  med_qrisk      <- round(median(dm_scored$qrisk10y, na.rm=TRUE), 1)
  pct_qrisk_gt10 <- .pct(dm_scored$qrisk10y > 10)
  pct_qrisk_gt20 <- .pct(dm_scored$qrisk10y > 20)
  mean_react     <- round(.mn(ascvd_scored$react5y), 1)
  med_react      <- round(median(ascvd_scored$react5y, na.rm=TRUE), 1)

  # ── Scenario risk shifts ──────────────────────────────────────────────────
  # At target: individual SBP and LDL gaps closed
  d_sbp_dm    <- pmax(dm_scored$sbp    - sbp_target,       0)
  d_sbp_asc   <- pmax(ascvd_scored$sbp - sbp_target,       0)
  d_ldl_dm    <- pmax(dm_scored$ldl    - ldl_target_dm,    0)
  d_ldl_asc   <- pmax(ascvd_scored$ldl - ldl_target_ascvd, 0)
  hr_dm_tgt   <- rr_sbp_dm^(d_sbp_dm/10)   * rr_ldl_dm^d_ldl_dm
  hr_asc_tgt  <- rr_sbp_ascvd^(d_sbp_asc/10) * rr_ldl_ascvd^d_ldl_asc
  qrisk_attarget <- round(.mn(clamp_vec(dm_scored$qrisk10y   * hr_dm_tgt,  0, 100)), 1)
  react_attarget <- round(.mn(clamp_vec(ascvd_scored$react5y * hr_asc_tgt, 0, 100)), 1)

  # 50% at target: each patient closes 50% of their SBP + LDL gap simultaneously
  hr_dm_50  <- rr_sbp_dm^(0.5*d_sbp_dm/10)    * rr_ldl_dm^(0.5*d_ldl_dm)
  hr_asc_50 <- rr_sbp_ascvd^(0.5*d_sbp_asc/10) * rr_ldl_ascvd^(0.5*d_ldl_asc)
  qrisk_50pct <- round(.mn(clamp_vec(dm_scored$qrisk10y   * hr_dm_50,  0, 100)), 1)
  react_50pct <- round(.mn(clamp_vec(ascvd_scored$react5y * hr_asc_50, 0, 100)), 1)
  delta_qrisk_50pct <- round(mean_qrisk  - qrisk_50pct, 1)
  delta_react_50pct <- round(mean_react  - react_50pct, 1)

  # 75% at target
  hr_dm_75  <- rr_sbp_dm^(0.75*d_sbp_dm/10)    * rr_ldl_dm^(0.75*d_ldl_dm)
  hr_asc_75 <- rr_sbp_ascvd^(0.75*d_sbp_asc/10) * rr_ldl_ascvd^(0.75*d_ldl_asc)
  qrisk_75pct <- round(.mn(clamp_vec(dm_scored$qrisk10y   * hr_dm_75,  0, 100)), 1)
  react_75pct <- round(.mn(clamp_vec(ascvd_scored$react5y * hr_asc_75, 0, 100)), 1)

  # 100% at target (individual gaps: same as qrisk_attarget / react_attarget)
  delta_qrisk_100pct <- round(mean_qrisk  - qrisk_attarget, 1)
  delta_react_100pct <- round(mean_react  - react_attarget, 1)

  # ── Treatment gap ─────────────────────────────────────────────────────────
  tg <- treatment_gap
  tg_dm  <- tg[tg$Cohort == "DM",  ]
  tg_asc <- tg[tg$Cohort == "ASCVD",]

  # ── Treatment intensity ───────────────────────────────────────────────────
  pct_no_llt_dm      <- .pct(dm_base$llt_int  == 0L)
  pct_high_llt_dm    <- .pct(dm_base$llt_int  == 3L)
  pct_max_llt_dm     <- .pct(dm_base$llt_int  == 4L)
  pct_no_llt_ascvd   <- .pct(ascvd_base$llt_int == 0L)
  pct_high_llt_ascvd <- .pct(ascvd_base$llt_int == 3L)
  pct_max_llt_ascvd  <- .pct(ascvd_base$llt_int == 4L)
  pct_no_aht_dm      <- .pct(dm_base$aht_int  == 0L)
  pct_no_aht_ascvd   <- .pct(ascvd_base$aht_int == 0L)

  # ── Sex-stratified treatment and target attainment (for Discussion) ───────
  .sx <- function(df, sx, expr)
    round(eval(substitute(expr), dplyr::filter(df, sex == sx)), 1)
  pct_statin_f_dm  <- .sx(dm_base,    "F", 100*mean(statin_rx_last6m==1L, na.rm=TRUE))
  pct_statin_m_dm  <- .sx(dm_base,    "M", 100*mean(statin_rx_last6m==1L, na.rm=TRUE))
  pct_statin_f_asc <- .sx(ascvd_base, "F", 100*mean(statin_rx_last6m==1L, na.rm=TRUE))
  pct_statin_m_asc <- .sx(ascvd_base, "M", 100*mean(statin_rx_last6m==1L, na.rm=TRUE))
  pct_ldl_f_dm     <- .sx(dm_base,    "F", 100*mean(ldl <= 2.0, na.rm=TRUE))
  pct_ldl_m_dm     <- .sx(dm_base,    "M", 100*mean(ldl <= 2.0, na.rm=TRUE))
  pct_ldl_f_asc    <- .sx(ascvd_base, "F", 100*mean(ldl <= 1.8, na.rm=TRUE))
  pct_ldl_m_asc    <- .sx(ascvd_base, "M", 100*mean(ldl <= 1.8, na.rm=TRUE))


  # ── WIMD summaries ────────────────────────────────────────────────────────
  wimd_labs <- c("1 (most\ndeprived)", "2", "3", "4", "5 (least\ndeprived)")
  dm_w  <- dm_scored    |> dplyr::filter(!is.na(wimd), wimd %in% 1:5) |>
            dplyr::mutate(wimd_f = factor(wimd, 1:5, wimd_labs))
  asc_w <- ascvd_scored |> dplyr::filter(!is.na(wimd), wimd %in% 1:5) |>
            dplyr::mutate(wimd_f = factor(wimd, 1:5, wimd_labs))

  w_dm <- dm_w  |> dplyr::group_by(wimd_f) |>
    dplyr::summarise(qrisk=.mn(qrisk10y), ldl_tgt=.pct(ldl<=2.0), .groups="drop")
  w_asc <- asc_w |> dplyr::group_by(wimd_f) |>
    dplyr::summarise(react=.mn(react5y),  ldl_tgt=.pct(ldl<=1.8), .groups="drop")
  w_pool <- dplyr::bind_rows(dm_w, asc_w) |> dplyr::group_by(wimd_f) |>
    dplyr::summarise(sbp_tgt=.pct(sbp<=130), statin=.pct(statin_rx_last6m==1L),
                     aht=.pct(treated_htn==1L), .groups="drop")

  .wv <- function(df, col, q) df[[col]][df$wimd_f == wimd_labs[q]]

  # ── Risk-quartile summaries ───────────────────────────────────────────────
  q_labs <- c("Q1 (lowest)", "Q2", "Q3", "Q4 (highest)")
  dm_q <- dm_scored |> dplyr::filter(!is.na(qrisk10y)) |>
    dplyr::mutate(risk_q = cut(qrisk10y, quantile(qrisk10y, c(0,.25,.5,.75,1), na.rm=TRUE),
                               labels=q_labs, include.lowest=TRUE))
  asc_q <- ascvd_scored |> dplyr::filter(!is.na(react5y)) |>
    dplyr::mutate(risk_q = cut(react5y, quantile(react5y, c(0,.25,.5,.75,1), na.rm=TRUE),
                               labels=q_labs, include.lowest=TRUE))
  .qv <- function(df, q, expr) {
    sub <- dplyr::filter(df, risk_q == q)
    round(eval(substitute(expr), sub), 1)
  }

  # ── Opportunity (Figure 3) ────────────────────────────────────────────────
  dm_q5  <- dm_q  |>
    dplyr::filter(!is.na(sbp), !is.na(ldl)) |>
    dplyr::mutate(
      risk5y = 1 - (1 - qrisk10y/100)^0.5,
      hr_tgt = rr_sbp_dm^(pmax(sbp - sbp_target, 0)/10) *
               rr_ldl_dm^pmax(ldl - ldl_target_dm, 0)
    )
  asc_q5 <- asc_q |>
    dplyr::filter(!is.na(sbp), !is.na(ldl)) |>
    dplyr::mutate(
      risk5y = react5y / 100,
      hr_tgt = rr_sbp_ascvd^(pmax(sbp - sbp_target, 0)/10) *
               rr_ldl_ascvd^pmax(ldl - ldl_target_ascvd, 0)
    )
  .opp_sq <- function(df, q)
    round(mean(dplyr::filter(df, risk_q==q)$risk5y, na.rm=TRUE) * 1000, 1)
  .opp_prev_cov <- function(df, q, cov, ldl_tgt, rr_sbp, rr_ldl) {
    s <- dplyr::filter(df, risk_q==q)
    hr_c <- rr_sbp^(cov*pmax(s$sbp-sbp_target,0)/10) * rr_ldl^(cov*pmax(s$ldl-ldl_tgt,0))
    round(mean(s$risk5y * (1 - hr_c), na.rm=TRUE) * 1000, 1)
  }
  .opp_prev_tgt <- function(df, q) {
    s <- dplyr::filter(df, risk_q==q)
    round(mean(s$risk5y * (1 - s$hr_tgt), na.rm=TRUE) * 1000, 1)
  }

  events_sq_dm_q1    <- .opp_sq(dm_q5,  "Q1 (lowest)")
  events_sq_dm_q4    <- .opp_sq(dm_q5,  "Q4 (highest)")
  events_sq_ascvd_q1 <- .opp_sq(asc_q5, "Q1 (lowest)")
  events_sq_ascvd_q4 <- .opp_sq(asc_q5, "Q4 (highest)")

  prevented_50pct_dm_q4    <- .opp_prev_cov(dm_q5,  "Q4 (highest)", 0.50, ldl_target_dm,    rr_sbp_dm,    rr_ldl_dm)
  prevented_50pct_dm_q1    <- .opp_prev_cov(dm_q5,  "Q1 (lowest)",  0.50, ldl_target_dm,    rr_sbp_dm,    rr_ldl_dm)
  prevented_50pct_ascvd_q4 <- .opp_prev_cov(asc_q5, "Q4 (highest)", 0.50, ldl_target_ascvd, rr_sbp_ascvd, rr_ldl_ascvd)
  prevented_50pct_ascvd_q1 <- .opp_prev_cov(asc_q5, "Q1 (lowest)",  0.50, ldl_target_ascvd, rr_sbp_ascvd, rr_ldl_ascvd)
  prevented_tgt_dm_q4    <- .opp_prev_tgt(dm_q5,  "Q4 (highest)")
  prevented_tgt_ascvd_q4 <- .opp_prev_tgt(asc_q5, "Q4 (highest)")

  # ── Assemble ──────────────────────────────────────────────────────────────
  list(
    n_total            = format(n_dm + n_ascvd, big.mark=","),
    n_dm               = format(n_dm, big.mark=","),
    n_ascvd            = format(n_ascvd, big.mark=","),
    mean_age_dm        = mean_age_dm,
    pct_male_dm        = pct_male_dm,
    mean_age_ascvd     = mean_age_ascvd,
    pct_male_ascvd     = pct_male_ascvd,
    pct_ihd            = pct_ihd,
    pct_stroke         = pct_stroke,
    pct_pad            = pct_pad,
    pct_htn_dm         = pct_htn_dm,
    pct_htn_ascvd      = pct_htn_ascvd,
    pct_hf_dm          = pct_hf_dm,
    pct_hf_ascvd       = pct_hf_ascvd,
    pct_wimd1_dm       = pct_wimd1_dm,
    pct_wimd1_ascvd    = pct_wimd1_ascvd,
    pct_statin_dm      = pct_statin_dm,
    pct_statin_ascvd   = pct_statin_ascvd,
    mean_ldl_dm        = mean_ldl_dm,
    sd_ldl_dm          = sd_ldl_dm,
    mean_ldl_ascvd     = mean_ldl_ascvd,
    sd_ldl_ascvd       = sd_ldl_ascvd,
    mean_sbp_dm        = mean_sbp_dm,
    sd_sbp_dm          = sd_sbp_dm,
    mean_sbp_ascvd     = mean_sbp_ascvd,
    sd_sbp_ascvd       = sd_sbp_ascvd,
    mean_qrisk         = mean_qrisk,
    med_qrisk          = med_qrisk,
    pct_qrisk_gt10     = pct_qrisk_gt10,
    pct_qrisk_gt20     = pct_qrisk_gt20,
    mean_react         = mean_react,
    med_react          = med_react,
    qrisk_50pct        = qrisk_50pct,
    react_50pct        = react_50pct,
    delta_qrisk_50pct  = delta_qrisk_50pct,
    delta_react_50pct  = delta_react_50pct,
    qrisk_75pct        = qrisk_75pct,
    react_75pct        = react_75pct,
    qrisk_attarget     = qrisk_attarget,
    react_attarget     = react_attarget,
    delta_qrisk_100pct = delta_qrisk_100pct,
    delta_react_100pct = delta_react_100pct,
    pct_ldl_tgt_dm     = tg_dm[["LDL at target (%)"]],
    pct_ldl_tgt_ascvd  = tg_asc[["LDL at target (%)"]],
    gap_ldl_dm         = tg_dm[["Mean LDL gap (mmol/L)"]],
    gap_ldl_ascvd      = tg_asc[["Mean LDL gap (mmol/L)"]],
    pct_sbp_tgt_dm     = tg_dm[["SBP at target (%)"]],
    pct_sbp_tgt_ascvd  = tg_asc[["SBP at target (%)"]],
    gap_sbp_dm         = tg_dm[["Mean SBP gap (mmHg)"]],
    gap_sbp_ascvd      = tg_asc[["Mean SBP gap (mmHg)"]],
    pct_no_llt_dm      = pct_no_llt_dm,
    pct_high_llt_dm    = pct_high_llt_dm,
    pct_max_llt_dm     = pct_max_llt_dm,
    pct_no_llt_ascvd   = pct_no_llt_ascvd,
    pct_high_llt_ascvd = pct_high_llt_ascvd,
    pct_max_llt_ascvd  = pct_max_llt_ascvd,
    pct_no_aht_dm      = pct_no_aht_dm,
    pct_no_aht_ascvd   = pct_no_aht_ascvd,
    pct_statin_f_dm    = pct_statin_f_dm,
    pct_statin_m_dm    = pct_statin_m_dm,
    pct_statin_f_asc   = pct_statin_f_asc,
    pct_statin_m_asc   = pct_statin_m_asc,
    pct_ldl_f_dm       = pct_ldl_f_dm,
    pct_ldl_m_dm       = pct_ldl_m_dm,
    pct_ldl_f_asc      = pct_ldl_f_asc,
    pct_ldl_m_asc      = pct_ldl_m_asc,
    events_sq_dm_q1        = events_sq_dm_q1,
    events_sq_dm_q4        = events_sq_dm_q4,
    events_sq_ascvd_q1     = events_sq_ascvd_q1,
    events_sq_ascvd_q4     = events_sq_ascvd_q4,
    prevented_50pct_dm_q4    = prevented_50pct_dm_q4,
    prevented_50pct_dm_q1    = prevented_50pct_dm_q1,
    prevented_50pct_ascvd_q4 = prevented_50pct_ascvd_q4,
    prevented_50pct_ascvd_q1 = prevented_50pct_ascvd_q1,
    prevented_tgt_dm_q4    = prevented_tgt_dm_q4,
    prevented_tgt_ascvd_q4 = prevented_tgt_ascvd_q4,
    qrisk_wimd1         = round(.wv(w_dm,  "qrisk",   1), 1),
    qrisk_wimd5         = round(.wv(w_dm,  "qrisk",   5), 1),
    react_wimd1         = round(.wv(w_asc, "react",   1), 1),
    react_wimd5         = round(.wv(w_asc, "react",   5), 1),
    pct_ldl_tgt_wimd1_dm    = .wv(w_dm,  "ldl_tgt", 1),
    pct_ldl_tgt_wimd5_dm    = .wv(w_dm,  "ldl_tgt", 5),
    pct_ldl_tgt_wimd1_ascvd = .wv(w_asc, "ldl_tgt", 1),
    pct_ldl_tgt_wimd5_ascvd = .wv(w_asc, "ldl_tgt", 5),
    pct_sbp_tgt_wimd1   = .wv(w_pool, "sbp_tgt", 1),
    pct_sbp_tgt_wimd5   = .wv(w_pool, "sbp_tgt", 5),
    pct_statin_wimd1    = .wv(w_pool, "statin",  1),
    pct_statin_wimd5    = .wv(w_pool, "statin",  5),
    pct_aht_wimd1       = .wv(w_pool, "aht",     1),
    pct_aht_wimd5       = .wv(w_pool, "aht",     5),
    pct_ldl_tgt_q4_dm   = .qv(dm_q,  "Q4 (highest)", 100*mean(ldl <= 2.0, na.rm=TRUE)),
    pct_ldl_tgt_q1_dm   = .qv(dm_q,  "Q1 (lowest)",  100*mean(ldl <= 2.0, na.rm=TRUE)),
    pct_sbp_tgt_q4_dm   = .qv(dm_q,  "Q4 (highest)", 100*mean(sbp <= 130, na.rm=TRUE)),
    pct_sbp_tgt_q1_dm   = .qv(dm_q,  "Q1 (lowest)",  100*mean(sbp <= 130, na.rm=TRUE)),
    pct_statin_q4_dm    = .qv(dm_q,  "Q4 (highest)", 100*mean(statin_rx_last6m==1L, na.rm=TRUE)),
    pct_statin_q4_ascvd = .qv(asc_q, "Q4 (highest)", 100*mean(statin_rx_last6m==1L, na.rm=TRUE)),
    pct_any_cand_dm     = round(100 * mean(
      (dm_scored$ldl  > ldl_target_dm    & dm_scored$llt_int < 3L) |
      (dm_scored$sbp  > sbp_target       & dm_scored$aht_int < 4L),
      na.rm = TRUE), 1),
    pct_any_cand_ascvd  = round(100 * mean(
      (ascvd_scored$ldl > ldl_target_ascvd & ascvd_scored$llt_int < 3L) |
      (ascvd_scored$sbp > sbp_target       & ascvd_scored$aht_int < 4L),
      na.rm = TRUE), 1)
  )
})
```


**Table 1.** Baseline characteristics

```{r table1, echo=FALSE}
kbl(baseline_table, align=c("l","c","c","c"),
    col.names=c("Characteristic","Overall","ASCVD","DM")) |>
  kable_styling(
    bootstrap_options=c("striped","hover","condensed"),
    full_width=FALSE, position="left"
  ) |>
  pack_rows("QRISK2-2015 risk factors", 4, 13,
            bold=TRUE, background="#F0F4FA", color="#2C3E6B") |>
  pack_rows("REACT risk factors", 14, 23,
            bold=TRUE, background="#F5F0FA", color="#4B2C6B") |>
  column_spec(1, width="18em") |>
  column_spec(2:4, width="8em")
```

<br>
<br>

```{r fig1, include=FALSE}
dm_plot_df    <- dm_scored    |> dplyr::filter(!is.na(qrisk10y))
ascvd_plot_df <- ascvd_scored |>
  dplyr::filter(!is.na(react5y)) |>
  dplyr::mutate(dx_type=dplyr::if_else(hard_event==1L,"MACE diagnosis","Stable diagnosis"))

dm_mean <- mean(dm_plot_df$qrisk10y);    dm_med <- median(dm_plot_df$qrisk10y)
ascvd_mean <- mean(ascvd_plot_df$react5y); ascvd_med <- median(ascvd_plot_df$react5y)

scen_pal_3 <- c("Status quo"      = "#e66914",
                "50% at target"  = "#bddd1d",
                "75% at target"  = "#34aa6f",
                "100% at target" = "#56026a")

p1a <- ggplot(dm_plot_df, aes(x=qrisk10y)) +
  geom_density(fill="#4D4D4D", alpha=0.35, linewidth=0.7) +
  geom_vline(xintercept=dm_med, linewidth=0.7, linetype="longdash") +
  annotate("text", x=Inf, y=Inf,
           label=sprintf("Mean %.1f%%\nMedian %.1f%%", dm_mean, dm_med),
           vjust=1.5, hjust=1.1, size=3.2) +
  labs(tag="A", title="DM (primary)",
       x="10-year risk (%)", y="Density") +
  theme_cap() + theme(plot.tag = element_text(face = "bold", size = 13))

dx_means <- ascvd_plot_df |>
  dplyr::group_by(dx_type) |>
  dplyr::summarise(mean_risk   = mean(react5y,   na.rm=TRUE),
                   median_risk = median(react5y, na.rm=TRUE),
                   .groups="drop") |>
  dplyr::mutate(dx_type   = factor(dx_type, levels=names(dx_pal)),
                lbl       = sprintf("Mean %.1f%%\nMedian %.1f%%", mean_risk, median_risk),
                vjust_val = (as.integer(factor(dx_type, levels=names(dx_pal))) - 1) * 2.5 + 1.5)

p1b <- ggplot(ascvd_plot_df, aes(x=react5y, colour=dx_type, fill=dx_type)) +
  geom_density(alpha=0.30, linewidth=0.7) +
  geom_vline(xintercept=ascvd_med, linewidth=0.7, linetype="longdash") +
  scale_colour_manual(values=dx_pal, name="Index diagnosis") +
  scale_fill_manual(values=dx_pal, name="Index diagnosis") +
  geom_text(data=dx_means,
            aes(x=Inf, y=Inf, label=lbl, colour=dx_type, vjust=vjust_val),
            hjust=1.1, size=3.0, lineheight=0.9, show.legend=FALSE) +
  labs(tag="B", title= "ASCVD (secondary)",
       x="5-year risk (%)", y="Density") +
  theme_cap() +
  theme(legend.position="right", legend.text=element_text(size=8),
  plot.tag = element_text(face = "bold", size = 13), axis.title.y = element_blank())

# ── Risk distributions under three scenarios ───────────────────────────────────
# Apply HR-based scenario shifts (no ABM) to produce overlaid distributions
rr_sbp_dm <- 0.90; rr_ldl_dm <- 0.79
rr_sbp_ascvd <- 0.80; rr_ldl_ascvd <- 0.78

apply_scenario <- function(dm_s, ascvd_s, cov=NULL) {
  dm_s2    <- dm_s
  ascvd_s2 <- ascvd_s
  if (is.null(cov)) {
    d_sbp_dm = d_sbp_ascvd = d_ldl_dm = d_ldl_ascvd = 0
  } else {
    d_sbp_dm    <- cov * pmax(dm_s$sbp    - sbp_target,       0)
    d_sbp_ascvd <- cov * pmax(ascvd_s$sbp - sbp_target,       0)
    d_ldl_dm    <- cov * pmax(dm_s$ldl    - ldl_target_dm,    0)
    d_ldl_ascvd <- cov * pmax(ascvd_s$ldl - ldl_target_ascvd, 0)
  }
  hr_dm    <- rr_sbp_dm^(d_sbp_dm/10)       * rr_ldl_dm^d_ldl_dm
  hr_ascvd <- rr_sbp_ascvd^(d_sbp_ascvd/10) * rr_ldl_ascvd^d_ldl_ascvd
  dm_s2$qrisk10y   <- clamp_vec(dm_s$qrisk10y   * hr_dm,    0, 100)
  ascvd_s2$react5y <- clamp_vec(ascvd_s$react5y * hr_ascvd, 0, 100)
  list(dm=dm_s2, ascvd=ascvd_s2)
}

risk_shift_df <- dplyr::bind_rows(
  { s <- apply_scenario(dm_scored, ascvd_scored)
    dplyr::bind_rows(
      dplyr::mutate(s$dm,    scenario="Status quo", cohort="DM"),
      dplyr::mutate(s$ascvd, scenario="Status quo", cohort="ASCVD"))},
  { s <- apply_scenario(dm_scored, ascvd_scored, cov=0.50)
    dplyr::bind_rows(
      dplyr::mutate(s$dm,    scenario="50% at target", cohort="DM"),
      dplyr::mutate(s$ascvd, scenario="50% at target", cohort="ASCVD"))},
  { s <- apply_scenario(dm_scored, ascvd_scored, cov=0.75)
    dplyr::bind_rows(
      dplyr::mutate(s$dm,    scenario="75% at target", cohort="DM"),
      dplyr::mutate(s$ascvd, scenario="75% at target", cohort="ASCVD"))},
  { s <- apply_scenario(dm_scored, ascvd_scored, cov=1.00)
    dplyr::bind_rows(
      dplyr::mutate(s$dm,    scenario="100% at target", cohort="DM"),
      dplyr::mutate(s$ascvd, scenario="100% at target", cohort="ASCVD"))}
)

scen_means_dm <- risk_shift_df |>
  dplyr::filter(cohort=="DM", !is.na(qrisk10y)) |>
  dplyr::group_by(scenario) |>
  dplyr::summarise(mean_risk   = mean(qrisk10y,   na.rm=TRUE),
                   median_risk = median(qrisk10y, na.rm=TRUE),
                   .groups="drop") |>
  dplyr::mutate(scenario  = factor(scenario, levels=names(scen_pal_3)),
                lbl       = sprintf("Mean %.1f%%\nMedian %.1f%%", mean_risk, median_risk),
                vjust_val = (as.integer(factor(scenario, levels=names(scen_pal_3))) - 1) * 2.5 + 1.5)

scen_means_ascvd <- risk_shift_df |>
  dplyr::filter(cohort=="ASCVD", !is.na(react5y)) |>
  dplyr::group_by(scenario) |>
  dplyr::summarise(mean_risk   = mean(react5y,   na.rm=TRUE),
                   median_risk = median(react5y, na.rm=TRUE),
                   .groups="drop") |>
  dplyr::mutate(scenario  = factor(scenario, levels=names(scen_pal_3)),
                lbl       = sprintf("Mean %.1f%%\nMedian %.1f%%", mean_risk, median_risk),
                vjust_val = (as.integer(factor(scenario, levels=names(scen_pal_3))) - 1) * 2.5 + 1.5)

p1c <- risk_shift_df |>
  dplyr::filter(cohort=="DM", !is.na(qrisk10y)) |>
  dplyr::mutate(scenario=factor(scenario, levels=names(scen_pal_3))) |>
  ggplot(aes(x=qrisk10y, colour=scenario, fill=scenario)) +
  geom_density(alpha=0.30, linewidth=0.7) +
  geom_vline(data=scen_means_dm,
             aes(xintercept=mean_risk, colour=scenario),
             linewidth=0.7, linetype="longdash", show.legend=FALSE) +
  geom_text(data=scen_means_dm,
            aes(x=Inf, y=Inf, label=lbl, colour=scenario, vjust=vjust_val),
            hjust=1.1, size=2.7, lineheight=0.9, show.legend=FALSE) +
  scale_colour_manual(values=scen_pal_3, name="Scenario") +
  scale_fill_manual(values=scen_pal_3, name="Scenario") +
  labs(tag= "A", x="10-year risk (%)", y="Density", title='DM (primary)') +
  theme_cap() +
  theme(legend.position="none", plot.tag = element_text(face = "bold", size = 13))

p1d <- risk_shift_df |>
  dplyr::filter(cohort=="ASCVD", !is.na(react5y)) |>
  dplyr::mutate(scenario=factor(scenario, levels=names(scen_pal_3))) |>
  ggplot(aes(x=react5y, colour=scenario, fill=scenario)) +
  geom_density(alpha=0.30, linewidth=0.7) +
  geom_vline(data=scen_means_ascvd,
             aes(xintercept=mean_risk, colour=scenario),
             linewidth=0.7, linetype="longdash", show.legend=FALSE) +
  geom_text(data=scen_means_ascvd,
            aes(x=Inf, y=Inf, label=lbl, colour=scenario, vjust=vjust_val),
            hjust=1.1, size=2.7, lineheight=0.9, show.legend=FALSE) +
  scale_colour_manual(values=scen_pal_3, name="Scenario") +
  scale_fill_manual(values=scen_pal_3, name="Scenario") +
  labs(tag= "B", x="5-year risk (%)", y="Density", title='ASCVD (secondary)') +
  theme_cap() +
  theme(legend.position="right", legend.text=element_text(size=8), plot.tag = element_text(face = "bold", size = 13), axis.title.y = element_blank())

```

<br>
<br>

```{r fig_biomarkers, fig.width=11, fig.asp=1.1, echo=FALSE}
# ── Shared data ────────────────────────────────────────────────────────────────
tg <- treatment_gap
cohort_pal <- c("DM (primary)"      = "#377EB8",
                "ASCVD (secondary)" = "#E41A1C")

biomarker_df <- dplyr::bind_rows(
  dm_base    |> dplyr::mutate(Cohort="DM (primary)"),
  ascvd_base |> dplyr::mutate(Cohort="ASCVD (secondary)")
) |>
  dplyr::mutate(Cohort=factor(Cohort, levels=names(cohort_pal)))

# Summary stats for in-panel annotations - one label per cohort per biomarker
ann_ldl <- tg |>
  dplyr::mutate(
    Cohort = dplyr::recode(Cohort,
      DM    = "DM (primary)",
      ASCVD = "ASCVD (secondary)"),
    label  = sprintf("At target: %s%%\nMean gap: %s mmol/L\nOn LLT: %s%%",
                     `LDL at target (%)`, `Mean LDL gap (mmol/L)`, `On LLT (%)`)
  ) |>
  dplyr::select(Cohort, label)

ann_sbp <- tg |>
  dplyr::mutate(
    Cohort = dplyr::recode(Cohort,
      DM    = "DM (primary)",
      ASCVD = "ASCVD (secondary)"),
    label  = sprintf("At target: %s%%\nMean gap: %s mmHg\nOn AHT: %s%%",
                     `SBP at target (%)`, `Mean SBP gap (mmHg)`, `On AHT (%)`)
  ) |>
  dplyr::select(Cohort, label)

# x positions for staggered annotation labels (avoid overlap)
ann_ldl$x_pos  <- c(6.5, 5.5)   # adjust if needed once rendered
ann_sbp$x_pos  <- c(185, 200)

# ── Panel A: LDL-C both cohorts on one plot ────────────────────────────────────
pBio_A <- ggplot(dplyr::filter(biomarker_df, !is.na(ldl)),
             aes(x=ldl, colour=Cohort, fill=Cohort)) +
  geom_density(alpha=0.22, linewidth=0.75) +
  geom_vline(xintercept=ldl_target_dm,    linetype="dashed", linewidth=0.7,
             colour=cohort_pal[["DM (primary)"]]) +
  geom_vline(xintercept=ldl_target_ascvd, linetype="dotted", linewidth=0.7,
             colour=cohort_pal[["ASCVD (secondary)"]]) +
  annotate("text", x=ldl_target_dm+0.07,    y=Inf, vjust=1.6, hjust=0,
           label="2.0 mmol/L",    size=2.5,
           colour=cohort_pal[["DM (primary)"]]) +
  annotate("text", x=ldl_target_ascvd-0.07, y=Inf, vjust=1.6, hjust=1,
           label="1.8 mmol/L",  size=2.5,
           colour=cohort_pal[["ASCVD (secondary)"]]) +
  scale_colour_manual(values=cohort_pal, name="Cohort") +
  scale_fill_manual(values=cohort_pal,   name="Cohort") +
  labs(tag="C", title='Lipids',
       x="LDL-C (mmol/L)", y="Density") +
  coord_cartesian(xlim = c(NA, 7.5)) +
  theme_cap() +
  theme(legend.position="none",
        plot.tag=element_text(face="bold", size=13))

# ── Panel B: SBP both cohorts on one plot ─────────────────────────────────────
pBio_B <- ggplot(dplyr::filter(biomarker_df, !is.na(sbp)),
             aes(x=sbp, colour=Cohort, fill=Cohort)) +
  geom_density(alpha=0.22, linewidth=0.75) +
  geom_vline(xintercept=sbp_target, linetype="dashed", linewidth=0.8,
             colour="#333333") +
  annotate("text", x=sbp_target+1, y=Inf, vjust=1.6, hjust=0,
           label="≤130 mmHg", size=2.5, colour="#333333") +
  scale_colour_manual(values=cohort_pal, name="Cohort") +
  scale_fill_manual(values=cohort_pal,   name="Cohort") +
  labs(tag="D", title='Blood pressure',
       x="SBP (mmHg)", y="Density") +
  theme_cap() +
  theme(legend.position="right", axis.title.y = element_blank(),
        plot.tag=element_text(face="bold", size=13))

# ── Panels C & D: treatment intensity bars ────────────────────────────────────
combined_lbl_map <- c(
  "No treatment"        = "No treatment",
  "Other treatment"     = "Other LLT / 1 class",
  "LI statin"           = "LI statin / 2 classes",
  "HI statin"           = "HI statin / 3 classes",
  "Combination therapy" = "Combination LLT / 4 classes",
  "1 drug class"        = "Other LLT / 1 class",
  "2 classes"           = "LI statin / 2 classes",
  "3 classes"           = "HI statin / 3 classes",
  "4 classes"           = "Combination LLT / 4 classes"
)
combined_pal <- c(
  "No treatment"                = "#e66914",
  "Other LLT / 1 class"         = "#bddd1d",
  "LI statin / 2 classes"       = "#34aa6f",
  "HI statin / 3 classes"       = "#4d5ca9",
  "Combination LLT / 4 classes" = "#56026a"
)

tr_clean <- treatment_rates |>
  dplyr::mutate(
    Label  = factor(
      dplyr::recode(as.character(Label), !!!combined_lbl_map),
      levels = rev(names(combined_pal))
    ),
    Cohort = factor(Cohort,
                    levels=c("DM (primary)",
                             "ASCVD (secondary)"))
  )

pBio_C <- ggplot(dplyr::filter(tr_clean, Treatment=="LLT"),
             aes(x=Cohort, y=Pct, fill=Label)) +
  geom_col(width=0.55, colour="white", linewidth=0.2) +
  scale_fill_manual(values=combined_pal, drop=FALSE) +
  scale_y_continuous(labels=scales::label_percent(scale=1), limits=c(0,100),
                     expand=expansion(mult=c(0,0.04))) +
  labs(tag="E", x='Cohort', y="Proportion") +
  theme_cap() +
  theme(axis.text.x=element_text(size=8.5),
        legend.text=element_text(size=8),
        legend.position="none",
        plot.tag=element_text(face="bold", size=13))

pBio_D <- ggplot(dplyr::filter(tr_clean, Treatment=="AHT"),
             aes(x=Cohort, y=Pct, fill=Label)) +
  geom_col(width=0.55, colour="white", linewidth=0.2) +
  scale_fill_manual(values=combined_pal, drop=FALSE, name="LLT / AHT") +
  scale_y_continuous(labels=scales::label_percent(scale=1), limits=c(0,100),
                     expand=expansion(mult=c(0,0.04))) +
  labs(tag="F", x='Cohort', y=NULL) +
  guides(fill=guide_legend(nrow=5)) +
  theme_cap() +
  theme(axis.text.x=element_text(size=8.5),
        legend.text=element_text(size=8),
        legend.position="right",
        plot.tag=element_text(face="bold", size=13))

# ── Compose 3×2 ───────────────────────────────────────────────────────────────
fig1 <- (p1a | p1b) / (pBio_A | pBio_B) / (pBio_C | pBio_D) +
  plot_layout(heights=c(1, 1, 1.05))
fig1
ggsave("figures/fig1.tiff", fig1, width = 11, height = 11 * 1.1, dpi = 300, compression = "lzw")
```

**Figure 1.** Estimated cardiovascular risk distributions and baseline treatment status. **(A)** QRISK2 10-year risk, DM (primary prevention). **(B)** REACT 5-year risk, ASCVD (secondary prevention). **(C)** LDL-C distributions; guideline targets (2.0 mmol/L DM, 1.8 mmol/L ASCVD). **(D)** SBP distributions; guideline target 130 mmHg. **(E)** Baseline lipid-lowering therapy. **(F)** Baseline antihypertensive therapy.


<br>
<br>


**Table 2.** Biomarker target scenarios - projected risk factor levels, target attainment, and cardiovascular risk by coverage level (Δ = absolute change from baseline; 95% CI from published relative risk bounds).

```{r table2, echo=FALSE}
tbl2 <- biomarker_table[, !names(biomarker_table) %in% "biomarker"]
bio_seq   <- rle(biomarker_table$biomarker)
start_idx <- cumsum(c(1, bio_seq$lengths[-length(bio_seq$lengths)]))
end_idx   <- cumsum(bio_seq$lengths)

k <- kbl(tbl2,
    digits=1, escape=FALSE,
    align=c("l","r","r","r","l","r","r","r","l"),
    col.names=c("Scenario",
                "Mean<br><small>(Δ from baseline)</small>",
                "At target (%)",
                "Projected QRISK2 (%)<br><small>(Δ from baseline)</small>",
                "95% CI",
                "Mean<br><small>(Δ from baseline)</small>",
                "At target (%)",
                "Projected REACT (%)<br><small>(Δ from baseline)</small>",
                "95% CI")) |>
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=FALSE, position="left") |>
  add_header_above(c(" "=1, "DM (primary)"=4, "ASCVD (secondary)"=4),
                   bold=TRUE, background="#F0F0F0") |>
  column_spec(1, width="12em") |>
  column_spec(c(2,3,6,7), width="8em") |>
  column_spec(4, width="8em", color="#2C3E6B", bold=TRUE) |>
  column_spec(5, width="7em", color="#2C3E6B", border_right=TRUE) |>
  column_spec(8, width="8em", color="#2C3E6B", bold=TRUE) |>
  column_spec(9, width="7em", color="#2C3E6B")

for (i in seq_along(bio_seq$values)) {
  bg <- if (bio_seq$values[i] == "LDL-C (mmol/L)") "#EEF4FB" else "#F5F0FA"
  col <- if (bio_seq$values[i] == "LDL-C (mmol/L)") "#2C3E6B" else "#4B2C6B"
  k <- pack_rows(k, bio_seq$values[i], start_idx[i], end_idx[i],
                 bold=TRUE, background=bg, color=col)
}
k
```

<br>
<br>

```{r fig_opportunity, fig.width=11, fig.asp=1.0, echo=FALSE}
# Shared quartile helpers (also used by Supplementary Figure 2)
q_base <- c("Q1 (lowest)","Q2","Q3","Q4 (highest)")
.q_range_labels <- function(brks, base = q_base) setNames(c(
  sprintf("%s\n(<%.1f%%)",       base[1], brks[2]),
  sprintf("%s\n(%.1f–%.1f%%)",   base[2], brks[2], brks[3]),
  sprintf("%s\n(%.1f–%.1f%%)",   base[3], brks[3], brks[4]),
  sprintf("%s\n(>%.1f%%)",       base[4], brks[4])
), base)

# ── Per-patient setup ────────────────────────────────────────────────────────
opp3_q_dm    <- quantile(dm_scored$qrisk10y,  c(0,.25,.5,.75,1), na.rm=TRUE)
opp3_q_ascvd <- quantile(ascvd_scored$react5y, c(0,.25,.5,.75,1), na.rm=TRUE)

dm_opp3 <- dm_scored |>
  dplyr::filter(!is.na(qrisk10y), !is.na(sbp), !is.na(ldl)) |>
  dplyr::mutate(
    risk5y = 1 - (1 - qrisk10y/100)^0.5,
    risk_q = cut(qrisk10y, breaks=opp3_q_dm, labels=q_base, include.lowest=TRUE)
  )

ascvd_opp3 <- ascvd_scored |>
  dplyr::filter(!is.na(react5y), !is.na(sbp), !is.na(ldl)) |>
  dplyr::mutate(
    risk5y = react5y / 100,
    risk_q = cut(react5y, breaks=opp3_q_ascvd, labels=q_base, include.lowest=TRUE)
  )

# ── RR point estimates and 95% CI bounds ──────────────────────────────────
# SBP: Ettehad 2016 (Lancet) subgroup estimates
# LDL: CTT diabetes 2008 (Lancet) for DM; CTT 2010 (Lancet) for ASCVD
rr_dm <- list(
  central = c(sbp=0.90, ldl=0.79),
  pessim  = c(sbp=0.95, ldl=0.86),  # upper 95% CI - smallest effect
  optimis = c(sbp=0.84, ldl=0.72)   # lower 95% CI - largest effect
)
rr_ascvd <- list(
  central = c(sbp=0.80, ldl=0.78),
  pessim  = c(sbp=0.83, ldl=0.80),
  optimis = c(sbp=0.77, ldl=0.76)
)

.make_opp3_rr <- function(df, cohort_nm, rr, ldl_tgt) {
  .ep_cov <- function(cov) {
    df |> dplyr::filter(!is.na(risk_q)) |>
      dplyr::mutate(hr_c = rr["sbp"]^(cov*pmax(sbp-sbp_target,0)/10) *
                           rr["ldl"]^(cov*pmax(ldl-ldl_tgt,0))) |>
      dplyr::group_by(risk_q) |>
      dplyr::summarise(events_expected  = mean(risk5y * hr_c, na.rm=TRUE) * 1000,
                       events_prevented = mean(risk5y * (1 - hr_c), na.rm=TRUE) * 1000,
                       .groups="drop")
  }
  dplyr::bind_rows(
    df |> dplyr::filter(!is.na(risk_q)) |>
      dplyr::group_by(risk_q) |>
      dplyr::summarise(events_expected=mean(risk5y,na.rm=TRUE)*1000,
                       events_prevented=0, .groups="drop") |>
      dplyr::mutate(scenario="Status quo"),
    .ep_cov(0.50) |> dplyr::mutate(scenario="50% at target"),
    .ep_cov(0.75) |> dplyr::mutate(scenario="75% at target"),
    .ep_cov(1.00) |> dplyr::mutate(scenario="100% at target")
  ) |> dplyr::mutate(Cohort = cohort_nm)
}

opp3_df <- dplyr::bind_rows(
  .make_opp3_rr(dm_opp3,    "DM (primary)",      rr_dm$central,    ldl_target_dm)    |> dplyr::mutate(bound="central"),
  .make_opp3_rr(dm_opp3,    "DM (primary)",      rr_dm$pessim,     ldl_target_dm)    |> dplyr::mutate(bound="pessim"),
  .make_opp3_rr(dm_opp3,    "DM (primary)",      rr_dm$optimis,    ldl_target_dm)    |> dplyr::mutate(bound="optimis"),
  .make_opp3_rr(ascvd_opp3, "ASCVD (secondary)", rr_ascvd$central, ldl_target_ascvd) |> dplyr::mutate(bound="central"),
  .make_opp3_rr(ascvd_opp3, "ASCVD (secondary)", rr_ascvd$pessim,  ldl_target_ascvd) |> dplyr::mutate(bound="pessim"),
  .make_opp3_rr(ascvd_opp3, "ASCVD (secondary)", rr_ascvd$optimis, ldl_target_ascvd) |> dplyr::mutate(bound="optimis")
) |>
  dplyr::mutate(
    scenario = factor(scenario, levels=names(scen_pal_3)),
    risk_q   = factor(risk_q,   levels=q_base),
    Cohort   = factor(Cohort,   levels=c("DM (primary)","ASCVD (secondary)"))
  )

# Quartile range labels
dm_qlbls3    <- .q_range_labels(opp3_q_dm)
ascvd_qlbls3 <- .q_range_labels(opp3_q_ascvd)

opp3_df <- opp3_df |>
  dplyr::mutate(
    risk_q_lbl = dplyr::if_else(
      Cohort == "DM (primary)",
      dm_qlbls3[as.character(risk_q)],
      ascvd_qlbls3[as.character(risk_q)]
    ),
    risk_q_lbl = factor(risk_q_lbl, levels=unique(c(dm_qlbls3, ascvd_qlbls3)))
  )

# Central estimate: used for bar panels, labels, and baseline
opp3_central <- opp3_df |> dplyr::filter(bound == "central")

opp3_base <- opp3_central |>
  dplyr::filter(scenario == "Status quo") |>
  dplyr::select(risk_q, Cohort, baseline = events_expected)

opp3_plot <- opp3_central |>
  dplyr::left_join(opp3_base, by=c("risk_q","Cohort"))

# RR uncertainty envelope for line panels (non-status-quo scenarios only)
# pessim  (upper CI of RR) → smallest effect → most events remain → ev_hi
# optimis (lower CI of RR) → largest effect  → fewest events remain → ev_lo
opp3_bounds_wide <- opp3_df |>
  dplyr::filter(bound != "central", scenario != "Status quo") |>
  dplyr::select(risk_q, risk_q_lbl, Cohort, scenario, bound, events_expected) |>
  tidyr::pivot_wider(names_from = bound, values_from = events_expected) |>
  dplyr::rename(ev_lo = optimis, ev_hi = pessim)

# ── Fixed y-axis limits (shared across cohort panels within each row) ─────
y_lim_line <- c(0, max(opp3_plot$baseline, na.rm = TRUE) * 1.08)
y_lim_bar  <- c(0, opp3_central |>
                     dplyr::filter(scenario == "100% at target") |>
                     dplyr::pull(events_prevented) |>
                     max(na.rm = TRUE) * 1.1)

# ── Panel function: line plots (A, B) ─────────────────────────────────────
.line_panel <- function(cohort_nm, q_lbls) {
  d <- opp3_plot |>
    dplyr::filter(Cohort == cohort_nm) |>
    dplyr::mutate(risk_q_lbl = factor(as.character(risk_q_lbl), levels = q_lbls))
  d_rib <- opp3_bounds_wide |>
    dplyr::filter(Cohort == cohort_nm) |>
    dplyr::mutate(risk_q_lbl = factor(as.character(risk_q_lbl), levels = q_lbls))
  ggplot(d, aes(x = risk_q_lbl, group = scenario)) +
    geom_ribbon(data = d_rib,
                aes(ymin = ev_lo, ymax = ev_hi, fill = scenario, group = scenario),
                alpha = 0.18, show.legend = FALSE) +
    geom_line(aes(y = events_expected, colour = scenario), linewidth = 1.1) +
    geom_point(aes(y = events_expected, colour = scenario, shape = scenario), size = 3.2) +
    scale_colour_manual(values = scen_pal_3, name = "Scenario") +
    scale_fill_manual(values = scen_pal_3) +
    scale_shape_manual(values = c("Status quo"=16, "50% at target"=17,
                                  "75% at target"=18, "100% at target"=15),
                       name = "Scenario") +
    scale_y_continuous(limits = y_lim_line, expand = expansion(mult = c(0, 0))) +
    labs(x = "Baseline risk quartile",
         y = "Events expected (per 1,000)",
         title = NULL) +
    theme_cap() +
    theme(legend.text = element_text(size = 9),
          axis.text.x = element_text(size = 8),
          axis.title.x = element_blank())
}

pOpp_A <- .line_panel("DM (primary)",      unname(dm_qlbls3)) + theme(legend.position = 'none') +
      labs(tag = "C") + theme(plot.tag = element_text(face = "bold", size = 13))
pOpp_B <- .line_panel("ASCVD (secondary)", unname(ascvd_qlbls3)) + theme(legend.position = 'right', axis.title.y = element_blank()) +
      labs(tag = "D") + theme(plot.tag = element_text(face = "bold", size = 13))

# ── Stacked bar data (C, D) ────────────────────────────────────────────────
bar_pal <- c("50% at target"               = "#bddd1d",
             "75% at target (additional)"  = "#34aa6f",
             "100% at target (additional)" = "#56026a")

opp3_stacked <- opp3_central |>
  dplyr::filter(scenario != "Status quo") |>
  dplyr::select(risk_q, risk_q_lbl, Cohort, scenario, events_prevented) |>
  tidyr::pivot_wider(names_from = scenario, values_from = events_prevented) |>
  dplyr::mutate(
    ep_50  = `50% at target`,
    ep_75m = `75% at target`  - `50% at target`,
    ep_100m= `100% at target` - `75% at target`
  ) |>
  dplyr::select(risk_q, risk_q_lbl, Cohort, ep_50, ep_75m, ep_100m) |>
  tidyr::pivot_longer(c(ep_50, ep_75m, ep_100m),
                      names_to = "component", values_to = "events") |>
  dplyr::mutate(component = factor(component,
    levels = c("ep_100m","ep_75m","ep_50"),
    labels = c("100% at target (additional)","75% at target (additional)","50% at target")))

# Error bars on total bar height (100% at target events prevented across RR bounds)
# pessim  → upper CI of RR → fewest events prevented → err_lo
# optimis → lower CI of RR → most events prevented  → err_hi
opp3_bar_err <- opp3_df |>
  dplyr::filter(scenario == "100% at target") |>
  dplyr::select(risk_q, risk_q_lbl, Cohort, bound, events_prevented) |>
  tidyr::pivot_wider(names_from = bound, values_from = events_prevented) |>
  dplyr::rename(err_lo = pessim, err_hi = optimis)

.bar_panel <- function(cohort_nm, q_lbls) {
  d <- opp3_stacked |>
    dplyr::filter(Cohort == cohort_nm) |>
    dplyr::mutate(risk_q_lbl = factor(as.character(risk_q_lbl), levels = q_lbls))
  d_err <- opp3_bar_err |>
    dplyr::filter(Cohort == cohort_nm) |>
    dplyr::mutate(risk_q_lbl = factor(as.character(risk_q_lbl), levels = q_lbls))
  ggplot(d, aes(x = risk_q_lbl, y = events, fill = component)) +
    geom_col(width = 0.6) +
    geom_errorbar(data = d_err,
                  aes(x = risk_q_lbl, ymin = err_lo, ymax = err_hi),
                  inherit.aes = FALSE,
                  width = 0.25, linewidth = 0.6, colour = "#e66914") +
    scale_fill_manual(values = bar_pal, name = "Scenario") +
    scale_y_continuous(limits = y_lim_bar, expand = expansion(mult = c(0, 0))) +
    labs(x = "Baseline risk quartile",
         y = "Events prevented (per 1,000)") +
    theme_cap() +
    theme(legend.text = element_text(size = 9),
          axis.text.x = element_text(size = 8))
}

pOpp_C <- .bar_panel("DM (primary)",      unname(dm_qlbls3)) + theme(legend.position = 'none') +
      labs(tag = "E") + theme(plot.tag = element_text(face = "bold", size = 13))
pOpp_D <- .bar_panel("ASCVD (secondary)", unname(ascvd_qlbls3)) + theme(legend.position = 'right', axis.title.y = element_blank()) +
      labs(tag = "F") + theme(plot.tag = element_text(face = "bold", size = 13))

# ── Compose 3 × 2 ─────────────────────────────────────────────────────────
fig2 <- (p1c | p1d) / (pOpp_A | pOpp_B) / (pOpp_C | pOpp_D) +
  plot_layout(heights=c(1, 1, 1.05))
fig2
ggsave("figures/fig2.tiff", fig2, width = 11, height = 11 * 1.0, dpi = 300, compression = "lzw")
```

**Figure 2.** Population-level cardiovascular risk distributions under modelled coverage scenarios and estimated event burden by risk quartile (individual patient gaps scaled by coverage). **(A)** QRISK2 distributions under scenarios. **(B)** REACT distributions under scenarios. **(C, D)** Expected five-year events per 1,000 patients by scenario (shaded ribbons denote 95% CI of published relative risk estimates). **(E, F)** Events prevented per 1,000 patients by scenario decomposed into incremental coverage gains (error bars denote 95% CI of published relative risk estimates for 100% at target).


<br>
<br>

```{r candidates_data, echo=FALSE}
# Quartile break points and range labels (mirrors .q_range_labels in fig_opportunity)
.cq_base   <- c("Q1 (lowest)", "Q2", "Q3", "Q4 (highest)")
.cq_brks_dm    <- quantile(dm_scored$qrisk10y,  c(0, .25, .5, .75, 1), na.rm = TRUE)
.cq_brks_ascvd <- quantile(ascvd_scored$react5y, c(0, .25, .5, .75, 1), na.rm = TRUE)

.mk_cq_lbls <- function(brks, base = .cq_base) setNames(c(
  sprintf("%s\n(<%.1f%%)",     base[1], brks[2]),
  sprintf("%s\n(%.1f–%.1f%%)", base[2], brks[2], brks[3]),
  sprintf("%s\n(%.1f–%.1f%%)", base[3], brks[3], brks[4]),
  sprintf("%s\n(>%.1f%%)",     base[4], brks[4])
), base)

cq_labs_dm    <- .mk_cq_lbls(.cq_brks_dm)
cq_labs_ascvd <- .mk_cq_lbls(.cq_brks_ascvd)

# Flag each patient by LDL and SBP treatment status
dm_cand <- dm_scored |>
  dplyr::filter(!is.na(qrisk10y), !is.na(ldl), !is.na(sbp)) |>
  dplyr::mutate(
    risk_q     = cut(qrisk10y, breaks = .cq_brks_dm,
                     labels = .cq_base, include.lowest = TRUE),
    risk_q_lbl = factor(cq_labs_dm[as.character(risk_q)],
                        levels = unname(cq_labs_dm)),
    risk5y = 1 - (1 - qrisk10y / 100)^0.5,
    ldl_status = dplyr::case_when(
      llt_int == 0     & ldl <= ldl_target_dm ~ "NC: No LLT, at target",
      llt_int %in% 1:2 & ldl <= ldl_target_dm ~ "NC: LI LLT, at target",
      llt_int >= 3     & ldl <= ldl_target_dm ~ "NC: HI LLT, at target",
      llt_int == 0     & ldl >  ldl_target_dm ~ "C: No LLT, not at target",
      llt_int %in% 1:2 & ldl >  ldl_target_dm ~ "C: LI LLT, not at target",
      llt_int >= 3     & ldl >  ldl_target_dm ~ "NC: HI LLT, not at target"
    ),
    sbp_status = dplyr::case_when(
      sbp <= sbp_target                    ~ "NC: At target",
      aht_int %in% 1:3 & sbp > sbp_target ~ "C: 1-3 agents, above target",
      aht_int == 4     & sbp > sbp_target ~ "NC: 4 agents, above target",
      aht_int == 0     & sbp > sbp_target ~ "C: No AHT, above target"
    ),
    ldl_cand = ldl > ldl_target_dm    & llt_int < 3,
    sbp_cand = sbp > sbp_target       & aht_int < 4,
    any_cand = ldl_cand | sbp_cand
  )

ascvd_cand <- ascvd_scored |>
  dplyr::filter(!is.na(react5y), !is.na(ldl), !is.na(sbp)) |>
  dplyr::mutate(
    risk_q     = cut(react5y, breaks = .cq_brks_ascvd,
                     labels = .cq_base, include.lowest = TRUE),
    risk_q_lbl = factor(cq_labs_ascvd[as.character(risk_q)],
                        levels = unname(cq_labs_ascvd)),
    risk5y = react5y / 100,
    ldl_status = dplyr::case_when(
      llt_int == 0     & ldl <= ldl_target_ascvd ~ "NC: No LLT, at target",
      llt_int %in% 1:2 & ldl <= ldl_target_ascvd ~ "NC: LI LLT, at target",
      llt_int >= 3     & ldl <= ldl_target_ascvd ~ "NC: HI LLT, at target",
      llt_int == 0     & ldl >  ldl_target_ascvd ~ "C: No LLT, not at target",
      llt_int %in% 1:2 & ldl >  ldl_target_ascvd ~ "C: LI LLT, not at target",
      llt_int >= 3     & ldl >  ldl_target_ascvd ~ "NC: HI LLT, not at target"
    ),
    sbp_status = dplyr::case_when(
      sbp <= sbp_target                    ~ "NC: At target",
      aht_int %in% 1:3 & sbp > sbp_target ~ "C: 1-3 agents, above target",
      aht_int == 4     & sbp > sbp_target ~ "NC: 4 agents, above target",
      aht_int == 0     & sbp > sbp_target ~ "C: No AHT, above target"
    ),
    ldl_cand = ldl > ldl_target_ascvd & llt_int < 3,
    sbp_cand = sbp > sbp_target       & aht_int < 4,
    any_cand = ldl_cand | sbp_cand
  )

# RR bounds - same values as fig_opportunity
.rr_cand <- list(
  DM    = list(central = c(sbp=0.90, ldl=0.79),
               pessim  = c(sbp=0.95, ldl=0.86),
               optimis = c(sbp=0.84, ldl=0.72)),
  ASCVD = list(central = c(sbp=0.80, ldl=0.78),
               pessim  = c(sbp=0.83, ldl=0.80),
               optimis = c(sbp=0.77, ldl=0.76))
)

# Events prevented per 1,000 candidates using patient-specific deltas to reach target
.cand_ep <- function(df, cohort, rr) {
  ldl_tgt <- if (cohort == "DM") ldl_target_dm else ldl_target_ascvd
  df |>
    dplyr::filter(any_cand, !is.na(risk_q)) |>
    dplyr::mutate(
      d_ldl = dplyr::if_else(ldl_cand, pmax(ldl - ldl_tgt, 0), 0),
      d_sbp = dplyr::if_else(sbp_cand, pmax(sbp - 130, 0), 0),
      hr    = rr["sbp"]^(d_sbp / 10) * rr["ldl"]^d_ldl,
      ep    = risk5y * (1 - hr)
    ) |>
    dplyr::group_by(risk_q, risk_q_lbl) |>
    dplyr::summarise(
      n_cand   = dplyr::n(),
      ep_per1k = mean(ep, na.rm = TRUE) * 1000,
      .groups  = "drop"
    ) |>
    dplyr::mutate(Cohort = if (cohort == "DM") "DM (primary)" else "ASCVD (secondary)")
}

# Compute all bounds and pivot to wide for error bars
cand_ep_wide <- dplyr::bind_rows(
  .cand_ep(dm_cand,    "DM",    .rr_cand$DM$central)    |> dplyr::mutate(bound = "central"),
  .cand_ep(dm_cand,    "DM",    .rr_cand$DM$pessim)     |> dplyr::mutate(bound = "pessim"),
  .cand_ep(dm_cand,    "DM",    .rr_cand$DM$optimis)    |> dplyr::mutate(bound = "optimis"),
  .cand_ep(ascvd_cand, "ASCVD", .rr_cand$ASCVD$central) |> dplyr::mutate(bound = "central"),
  .cand_ep(ascvd_cand, "ASCVD", .rr_cand$ASCVD$pessim)  |> dplyr::mutate(bound = "pessim"),
  .cand_ep(ascvd_cand, "ASCVD", .rr_cand$ASCVD$optimis) |> dplyr::mutate(bound = "optimis")
) |>
  tidyr::pivot_wider(
    id_cols     = c(risk_q, risk_q_lbl, Cohort, n_cand),
    names_from  = bound,
    values_from = ep_per1k
  ) |>
  dplyr::rename(ep_per1k = central, err_lo = pessim, err_hi = optimis)

cand_ep_dm    <- dplyr::filter(cand_ep_wide, Cohort == "DM (primary)")
cand_ep_ascvd <- dplyr::filter(cand_ep_wide, Cohort == "ASCVD (secondary)")

# Scenario 1: only treatment-naïve candidates (llt_int == 0 / aht_int == 0) are
# initiated on high-intensity statin (~1.0 mmol/L LDL reduction) and standard
# first-line antihypertensive (~10 mmHg SBP reduction), capped at patient gap.
# Undertreated patients contribute zero delta in this scenario.
.cand_ep_initiated <- function(df, cohort, rr,
                                d_ldl_init = 1.0, d_sbp_init = 10) {
  ldl_tgt <- if (cohort == "DM") ldl_target_dm else ldl_target_ascvd
  df |>
    dplyr::filter(any_cand, !is.na(risk_q)) |>
    dplyr::mutate(
      d_ldl = dplyr::if_else(ldl_cand & llt_int == 0,
                             pmin(pmax(ldl - ldl_tgt, 0), d_ldl_init), 0),
      d_sbp = dplyr::if_else(sbp_cand & aht_int == 0,
                             pmin(pmax(sbp - 130,    0), d_sbp_init), 0),
      hr    = rr["sbp"]^(d_sbp / 10) * rr["ldl"]^d_ldl,
      ep    = risk5y * (1 - hr)
    ) |>
    dplyr::group_by(risk_q, risk_q_lbl) |>
    dplyr::summarise(ep_per1k = mean(ep, na.rm = TRUE) * 1000, .groups = "drop") |>
    dplyr::mutate(Cohort = if (cohort == "DM") "DM (primary)" else "ASCVD (secondary)")
}

cand_scen <- dplyr::bind_rows(
  .cand_ep_initiated(dm_cand,    "DM",    .rr_cand$DM$central)    |>
    dplyr::mutate(Scenario = "All candidates treated"),
  .cand_ep_initiated(ascvd_cand, "ASCVD", .rr_cand$ASCVD$central) |>
    dplyr::mutate(Scenario = "All candidates treated"),
  cand_ep_wide |>
    dplyr::select(risk_q, risk_q_lbl, Cohort, ep_per1k) |>
    dplyr::mutate(Scenario = "Candidates reach target")
) |>
  dplyr::mutate(Scenario = factor(Scenario,
    levels = c("All candidates treated", "Candidates reach target")))

cand_scen_pal <- c(
  "All candidates treated" = "#bddd1d",
  "Candidates reach target"    = "#34aa6f"
)

cand_scen_dm    <- dplyr::filter(cand_scen, Cohort == "DM (primary)")
cand_scen_ascvd <- dplyr::filter(cand_scen, Cohort == "ASCVD (secondary)")

# Stacked bar data for E/F: bottom = "All candidates treated", top = marginal
# Mirrors figure 3 C/D stacking logic
cand_stacked <- cand_scen |>
  dplyr::select(risk_q, risk_q_lbl, Cohort, ep_per1k, Scenario) |>
  tidyr::pivot_wider(names_from = Scenario, values_from = ep_per1k) |>
  dplyr::mutate(marginal_ep = `Candidates reach target` - `All candidates treated`) |>
  dplyr::select(-`Candidates reach target`) |>
  dplyr::rename(initiated_ep = `All candidates treated`) |>
  tidyr::pivot_longer(c(initiated_ep, marginal_ep),
                      names_to = "component", values_to = "ep_per1k") |>
  dplyr::mutate(component = factor(component,
    levels = c("marginal_ep", "initiated_ep"),
    labels = c("Additional to target", "All candidates treated")))

cand_stacked_dm    <- dplyr::filter(cand_stacked, Cohort == "DM (primary)")
cand_stacked_ascvd <- dplyr::filter(cand_stacked, Cohort == "ASCVD (secondary)")

cand_bar_pal <- c(
  "All candidates treated" = "#bddd1d",
  "Additional to target"   = "#56026a"
)

# CI bounds for "All candidates treated" (treatment-naive initiated only)
cand_init_wide <- dplyr::bind_rows(
  .cand_ep_initiated(dm_cand,    "DM",    .rr_cand$DM$pessim)    |> dplyr::mutate(bound = "pessim"),
  .cand_ep_initiated(dm_cand,    "DM",    .rr_cand$DM$optimis)   |> dplyr::mutate(bound = "optimis"),
  .cand_ep_initiated(ascvd_cand, "ASCVD", .rr_cand$ASCVD$pessim) |> dplyr::mutate(bound = "pessim"),
  .cand_ep_initiated(ascvd_cand, "ASCVD", .rr_cand$ASCVD$optimis)|> dplyr::mutate(bound = "optimis")
) |>
  tidyr::pivot_wider(id_cols = c(risk_q, risk_q_lbl, Cohort),
                     names_from = bound, values_from = ep_per1k) |>
  dplyr::rename(err_lo = pessim, err_hi = optimis)
cand_init_dm    <- dplyr::filter(cand_init_wide, Cohort == "DM (primary)")
cand_init_ascvd <- dplyr::filter(cand_init_wide, Cohort == "ASCVD (secondary)")

# Inline scalars
.p2 <- function(x) round(100 * mean(x, na.rm = TRUE), 1)

pct_any_cand_dm       <- .p2(dm_cand$any_cand)
pct_ldl_cand_dm       <- .p2(dm_cand$ldl_cand)
pct_sbp_cand_dm       <- .p2(dm_cand$sbp_cand)
pct_any_cand_ascvd    <- .p2(ascvd_cand$any_cand)
pct_ldl_cand_ascvd    <- .p2(ascvd_cand$ldl_cand)
pct_sbp_cand_ascvd    <- .p2(ascvd_cand$sbp_cand)

ep_cand_dm_q1    <- round(cand_ep_dm$ep_per1k[cand_ep_dm$risk_q == "Q1 (lowest)"], 1)
ep_cand_dm_q4    <- round(cand_ep_dm$ep_per1k[cand_ep_dm$risk_q == "Q4 (highest)"], 1)
ep_cand_ascvd_q1 <- round(cand_ep_ascvd$ep_per1k[cand_ep_ascvd$risk_q == "Q1 (lowest)"], 1)
ep_cand_ascvd_q4 <- round(cand_ep_ascvd$ep_per1k[cand_ep_ascvd$risk_q == "Q4 (highest)"], 1)
```

```{r fig_candidates, fig.width=8, fig.asp=1, echo=FALSE}
ldl_pal    <- c("NC: No LLT, at target"     = "#a8d5b8",
                "NC: LI LLT, at target"     = "#d5edcf",
                "NC: HI LLT, at target"     = "#b8d8eb",
                "NC: HI LLT, not at target" = "#8fafc8",
                "C: LI LLT, not at target"     = "#fdae61",
                "C: No LLT, not at target"     = "#d73027")
ldl_levels <- c("NC: HI LLT, at target","NC: HI LLT, not at target",
                "NC: No LLT, at target","NC: LI LLT, at target",
                "C: LI LLT, not at target","C: No LLT, not at target")

sbp_pal    <- c("NC: At target"               = "#a8d5b8",
                "C: 1-3 agents, above target"    = "#fdae61",
                "NC: 4 agents, above target"  = "#8fafc8",
                "C: No AHT, above target"        = "#d73027")
sbp_levels <- c("NC: 4 agents, above target","NC: At target",
                "C: 1-3 agents, above target","C: No AHT, above target")
.cq_col4    <- c("#2171B5", "#6BAED6", "#F16913", "#CB181D")

# Build a named colour palette from a labelled quartile vector
.cq_pal_named <- function(q_lbls) setNames(.cq_col4, unname(q_lbls))

.ldl_status_q <- function(df, cohort_nm) {
  df |>
    dplyr::filter(!is.na(risk_q_lbl), !is.na(ldl_status)) |>
    dplyr::mutate(ldl_status = factor(ldl_status, levels = ldl_levels)) |>
    dplyr::count(risk_q_lbl, ldl_status, .drop = FALSE) |>
    dplyr::group_by(risk_q_lbl) |>
    dplyr::mutate(pct = 100 * n / sum(n)) |>
    dplyr::ungroup() |>
    dplyr::mutate(Cohort = cohort_nm)
}

ldl_status_dm    <- .ldl_status_q(dm_cand,    "DM (primary)")
ldl_status_ascvd <- .ldl_status_q(ascvd_cand, "ASCVD (secondary)")

.sbp_status_q <- function(df, cohort_nm) {
  df |>
    dplyr::filter(!is.na(risk_q_lbl), !is.na(sbp_status)) |>
    dplyr::mutate(sbp_status = factor(sbp_status, levels = sbp_levels)) |>
    dplyr::count(risk_q_lbl, sbp_status, .drop = FALSE) |>
    dplyr::group_by(risk_q_lbl) |>
    dplyr::mutate(pct = 100 * n / sum(n)) |>
    dplyr::ungroup() |>
    dplyr::mutate(Cohort = cohort_nm)
}

sbp_status_dm    <- .sbp_status_q(dm_cand,    "DM (primary)")
sbp_status_ascvd <- .sbp_status_q(ascvd_cand, "ASCVD (secondary)")

.ldl_bar <- function(d, q_lbls, tag, title_nm, show_legend = FALSE) {
  p <- ggplot(d, aes(x = risk_q_lbl, y = pct, fill = ldl_status)) +
    geom_col(width = 0.7) +
    scale_fill_manual(values = ldl_pal, drop = FALSE) +
    scale_x_discrete(limits = unname(q_lbls)) +
    scale_y_continuous(limits = c(0, 101), expand = expansion(mult = c(0, 0))) +
    labs(tag = tag, title = title_nm, x = NULL, y = "Proportion (%)") +
    theme_cap() +
    theme(plot.tag = element_text(face = "bold", size = 13),
          axis.text.x = element_text(size = 7.5))
  if (!show_legend) p + theme(legend.position = "none") else p
}

.sbp_bar <- function(d, q_lbls, tag, title_nm, show_legend = FALSE) {
  p <- ggplot(d, aes(x = risk_q_lbl, y = pct, fill = sbp_status)) +
    geom_col(width = 0.7) +
    scale_fill_manual(values = sbp_pal, drop = FALSE) +
    scale_x_discrete(limits = unname(q_lbls)) +
    scale_y_continuous(limits = c(0, 101), expand = expansion(mult = c(0, 0))) +
    labs(tag = tag, title = title_nm, x = NULL, y = "Proportion (%)") +
    theme_cap() +
    theme(plot.tag = element_text(face = "bold", size = 13),
          axis.text.x = element_text(size = 7.5))
  if (!show_legend) p + theme(legend.position = "none") else p
}

pC_A <- .ldl_bar(ldl_status_dm,    cq_labs_dm,    "A", "DM (primary)")
pC_B <- .ldl_bar(ldl_status_ascvd, cq_labs_ascvd, "B", "ASCVD (secondary)", show_legend = TRUE) +
  theme(axis.title.y = element_blank(), legend.position = "right") +
  labs(fill = 'LDL-C')
pC_C <- .sbp_bar(sbp_status_dm,    cq_labs_dm,    "C", NULL)
pC_D <- .sbp_bar(sbp_status_ascvd, cq_labs_ascvd, "D", NULL, show_legend = TRUE) +
  theme(axis.title.y = element_blank(), legend.position = "right") +
  labs(fill = 'Blood pressure')

# Events prevented per 1,000 candidates: stacked bars + error bars matching fig3 C/D
# Bottom = "All candidates treated" (untreated initiated), top = marginal to target
# Error bars span 95% CI of total "Candidates reach target" estimate
.ep_bar <- function(stacked_df, ep_df, q_lbls, tag, show_y = TRUE, init_df = NULL) {
  ymax <- ceiling(max(ep_df$err_hi, na.rm = TRUE) * 1.15)

  d <- stacked_df |>
    dplyr::mutate(risk_q_lbl = factor(risk_q_lbl, levels = unname(q_lbls)))
  d_err <- ep_df |>
    dplyr::mutate(risk_q_lbl = factor(risk_q_lbl, levels = unname(q_lbls)))

  p <- ggplot(d, aes(x = risk_q_lbl, y = ep_per1k, fill = component)) +
    geom_col(width = 0.6) +
    geom_errorbar(data = d_err,
                  aes(x = risk_q_lbl, ymin = err_lo, ymax = err_hi),
                  inherit.aes = FALSE,
                  width = 0.25, linewidth = 0.6, colour = "#e66914") +
    scale_fill_manual(values = cand_bar_pal, name = "Scenario") +
    scale_x_discrete(limits = unname(q_lbls)) +
    scale_y_continuous(limits = c(0, ymax), expand = expansion(mult = c(0, 0))) +
    labs(tag = tag,
         x = "Baseline risk quartile",
         y = "Events prevented (per 1,000)") +
    theme_cap() +
    theme(plot.tag = element_text(face = "bold", size = 13),
          axis.text.x = element_text(size = 7.5))

  if (!is.null(init_df)) {
    d_init <- init_df |>
      dplyr::mutate(risk_q_lbl = factor(risk_q_lbl, levels = unname(q_lbls)))
    p <- p + geom_errorbar(data = d_init,
                           aes(x = risk_q_lbl, ymin = err_lo, ymax = err_hi),
                           inherit.aes = FALSE,
                           width = 0.18, linewidth = 0.6, colour = "#e66914")
  }

  if (!show_y) p + theme(axis.title.y = element_blank(), legend.position = "none")
  else p + theme(legend.position = "right",
                 legend.text = element_text(size = 8),
                 legend.key.height = unit(2, "lines"))
}

pC_E <- .ep_bar(cand_stacked_dm,    cand_ep_dm,    cq_labs_dm,    "E", show_y = TRUE)  + theme(legend.position = "none")
pC_F <- .ep_bar(cand_stacked_ascvd, cand_ep_ascvd, cq_labs_ascvd, "F", show_y = FALSE) + theme(legend.position = "right")

fig3 <- (pC_A | pC_B) / (pC_C | pC_D) / (pC_E | pC_F) +
  plot_layout(heights = c(1, 1, 1))
fig3
ggsave("figures/fig3.tiff", fig3, width = 11, height = 11 * 1.1, dpi = 300, compression = "lzw")
```

**Figure 3.** LDL-C and blood pressure treatment status by baseline risk quartile and projected prevention opportunity. **(A, B)** LDL-C treatment status by LLT use and target attainment. **(C, D)** SBP treatment status by antihypertensive agent count and target attainment. **(E, F)** Projected five-year events prevented per 1,000 candidates using patient-specific risk factor gaps (error bars denote 95% CI of published relative risk estimates).


<br>
<br>

```{r interact_data, echo=FALSE}
q_ix_dm    <- quantile(dm_scored$qrisk10y,   probs = c(0,.25,.5,.75,1), na.rm = TRUE)
q_ix_ascvd <- quantile(ascvd_scored$react5y, probs = c(0,.25,.5,.75,1), na.rm = TRUE)
q_labs_ix  <- c("Q1 (lowest)","Q2","Q3","Q4 (highest)")
wimd_labs_ix <- c("1 (most\ndeprived)","2","3","4","5 (least\ndeprived)")

ix_dm <- dm_scored |>
  dplyr::filter(!is.na(wimd), wimd %in% 1:5, !is.na(qrisk10y)) |>
  dplyr::mutate(
    risk_q = cut(qrisk10y, breaks = q_ix_dm,    labels = q_labs_ix, include.lowest = TRUE),
    wimd_f = factor(wimd, levels = 1:5, labels = wimd_labs_ix)
  )
ix_ascvd <- ascvd_scored |>
  dplyr::filter(!is.na(wimd), wimd %in% 1:5, !is.na(react5y)) |>
  dplyr::mutate(
    risk_q = cut(react5y, breaks = q_ix_ascvd, labels = q_labs_ix, include.lowest = TRUE),
    wimd_f = factor(wimd, levels = 1:5, labels = wimd_labs_ix)
  )

ix_sum <- dplyr::bind_rows(
  ix_dm |>
    dplyr::filter(!is.na(risk_q)) |>
    dplyr::group_by(wimd_f, risk_q) |>
    dplyr::summarise(
      pct_ldl_target       = 100 * mean(ldl <= 2.0,                     na.rm = TRUE),
      pct_on_statin        = 100 * mean(statin_rx_last6m == 1L,         na.rm = TRUE),
      pct_sbp_target       = 100 * mean(sbp <= 130,                     na.rm = TRUE),
      pct_on_aht           = 100 * mean(treated_htn == 1L,              na.rm = TRUE),
      pct_ldl_untreated    = 100 * mean(ldl > 2.0 & llt_int == 0,       na.rm = TRUE),
      pct_ldl_undertreated = 100 * mean(ldl > 2.0 & llt_int %in% 1:3,  na.rm = TRUE),
      pct_sbp_untreated    = 100 * mean(sbp > 130  & aht_int == 0,      na.rm = TRUE),
      pct_sbp_undertreated = 100 * mean(sbp > 130  & aht_int %in% 1:3, na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::mutate(Cohort = "DM (primary)"),
  ix_ascvd |>
    dplyr::filter(!is.na(risk_q)) |>
    dplyr::group_by(wimd_f, risk_q) |>
    dplyr::summarise(
      pct_ldl_target       = 100 * mean(ldl <= 1.8,                     na.rm = TRUE),
      pct_on_statin        = 100 * mean(statin_rx_last6m == 1L,         na.rm = TRUE),
      pct_sbp_target       = 100 * mean(sbp <= 130,                     na.rm = TRUE),
      pct_on_aht           = 100 * mean(treated_htn == 1L,              na.rm = TRUE),
      pct_ldl_untreated    = 100 * mean(ldl > 1.8 & llt_int == 0,       na.rm = TRUE),
      pct_ldl_undertreated = 100 * mean(ldl > 1.8 & llt_int %in% 1:3,  na.rm = TRUE),
      pct_sbp_untreated    = 100 * mean(sbp > 130  & aht_int == 0,      na.rm = TRUE),
      pct_sbp_undertreated = 100 * mean(sbp > 130  & aht_int %in% 1:3, na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::mutate(Cohort = "ASCVD (secondary)")
)

.ixv <- function(co, wq, rq, mt)
  round(ix_sum[[mt]][ix_sum$Cohort == co &
                     as.character(ix_sum$wimd_f) == wq &
                     as.character(ix_sum$risk_q) == rq], 1)
pct_ldl_q4w1_dm  <- .ixv("DM (primary)",      "1 (most\ndeprived)",  "Q4 (highest)", "pct_ldl_target")
pct_ldl_q4w5_dm  <- .ixv("DM (primary)",      "5 (least\ndeprived)", "Q4 (highest)", "pct_ldl_target")
pct_ldl_q4w1_asc <- .ixv("ASCVD (secondary)", "1 (most\ndeprived)",  "Q4 (highest)", "pct_ldl_target")
pct_ldl_q4w5_asc <- .ixv("ASCVD (secondary)", "5 (least\ndeprived)", "Q4 (highest)", "pct_ldl_target")
```

```{r fig_interact, include=FALSE}
# ── Colour limits: data range across both cohorts per metric, with 2 pp padding
ldl_lims <- range(ix_sum$pct_ldl_target, na.rm = TRUE) + c(-2, 2)
stn_lims <- range(ix_sum$pct_on_statin,  na.rm = TRUE) + c(-2, 2)
sbp_lims <- range(ix_sum$pct_sbp_target, na.rm = TRUE) + c(-2, 2)
aht_lims <- range(ix_sum$pct_on_aht,     na.rm = TRUE) + c(-2, 2)

# ── Heatmap panel function ─────────────────────────────────────────────────
.heat_panel <- function(dat, cohort_nm, metric_col, fill_label, tag,
                        limits, show_legend = FALSE) {
  d <- dat |>
    dplyr::filter(Cohort == cohort_nm, !is.na(risk_q)) |>
    dplyr::mutate(risk_q = factor(as.character(risk_q), levels = rev(q_labs_ix)))
  p <- ggplot(d, aes(x = wimd_f, y = risk_q, fill = .data[[metric_col]])) +
    geom_tile(colour = "white", linewidth = 0.6) +
    geom_text(aes(label = sprintf("%.0f%%", .data[[metric_col]])),
              size = 2.8, colour = "white", fontface = "bold") +
    scale_fill_gradient(low = "#CB181D", high = "#2171B5",
                        limits = limits, name = fill_label,
                        labels = scales::label_percent(scale = 1)) +
    labs(tag = tag, title = NULL,
         x = NULL, y = NULL) +
    theme_cap() +
    theme(plot.tag    = element_text(face = "bold", size = 13),
          axis.text.x = element_text(size = 8))
  if (show_legend) p + theme(legend.position = "right")
  else             p + theme(legend.position = "none")
}

pIA <- .heat_panel(ix_sum, "DM (primary)",      "pct_ldl_target", "LDL-C\nat target (%)", "A", ldl_lims) + labs(title='DM (primary)', y = "Baseline risk quartile") + theme(axis.text.x = element_blank())
pIB <- .heat_panel(ix_sum, "ASCVD (secondary)", "pct_ldl_target", "LDL-C\nat target (%)", "B", ldl_lims, show_legend = TRUE) + labs(title='ASCVD (secondary)') + theme(axis.text.x = element_blank(), axis.text.y = element_blank())
pIC <- .heat_panel(ix_sum, "DM (primary)",      "pct_on_statin",  "On statin\n(%)",        "C", stn_lims) + labs(y = "Baseline risk quartile") + theme(axis.text.x = element_blank())
pID <- .heat_panel(ix_sum, "ASCVD (secondary)", "pct_on_statin",  "On statin\n(%)",        "D", stn_lims, show_legend = TRUE) + theme(axis.text.x = element_blank(), axis.text.y = element_blank())
pIE <- .heat_panel(ix_sum, "DM (primary)",      "pct_sbp_target", "SBP\nat target (%)",    "E", sbp_lims) + labs(y = "Baseline risk quartile") + theme(axis.text.x = element_blank())
pIF <- .heat_panel(ix_sum, "ASCVD (secondary)", "pct_sbp_target", "SBP\nat target (%)",    "F", sbp_lims, show_legend = TRUE) + theme(axis.text.x = element_blank(), axis.text.y = element_blank())
pIG <- .heat_panel(ix_sum, "DM (primary)",      "pct_on_aht",     "On AHT\n(%)",           "G", aht_lims) + labs(y = "Baseline risk quartile", x = "WIMD quintile")
pIH <- .heat_panel(ix_sum, "ASCVD (secondary)", "pct_on_aht",     "On AHT\n(%)",           "H", aht_lims, show_legend = TRUE) + labs(x = "WIMD quintile") + theme(axis.text.y = element_blank())

fig8 <- (pIA | pIB) / (pIC | pID) / (pIE | pIF) / (pIG | pIH)
fig8
ggsave("figures/fig8.tiff", fig8, width = 11, height = 11 * 1.65, dpi = 300, compression = "lzw")
```

<br>
<br>

```{r suppl_table1, echo=FALSE}
# Classify candidates as untreated vs undertreated
# Untreated: at least one contributing condition has intensity 0
# Undertreated: all contributing conditions have intensity 1-3
dm_cand_tbl <- dm_cand |>
  dplyr::filter(any_cand) |>
  dplyr::mutate(
    tx_status = dplyr::if_else(
      (ldl_cand & llt_int == 0) | (sbp_cand & aht_int == 0),
      "Untreated", "Undertreated"
    )
  )

ascvd_cand_tbl <- ascvd_cand |>
  dplyr::filter(any_cand) |>
  dplyr::mutate(
    tx_status = dplyr::if_else(
      (ldl_cand & llt_int == 0) | (sbp_cand & aht_int == 0),
      "Untreated", "Undertreated"
    )
  )

pct  <- function(x) sprintf("%d (%.1f%%)", sum(x, na.rm = TRUE), 100 * mean(x, na.rm = TRUE))
mn   <- function(x) sprintf("%.1f (%.1f)",  mean(x, na.rm = TRUE), sd(x, na.rm = TRUE))
mn2  <- function(x) sprintf("%.2f (%.2f)", mean(x, na.rm = TRUE), sd(x, na.rm = TRUE))

summ_cand <- function(df, is_dm = TRUE) {
  vt <- function(val) if (is_dm) "-" else pct(df$vascular_territory == val)
  list(
    "N"                               = nrow(df),
    "Age"                             = mn(df$age),
    "Male"                            = pct(df$sex == "M"),
    "Statin in last 6m"               = pct(df$statin_rx_last6m),
    "Systolic blood pressure"         = mn(df$sbp),
    "LDL-C"                           = mn2(df$ldl),
    "Body Mass Index"                 = mn(df$bmi),
    "WIMD quintile 1 (most deprived)" = pct(df$wimd == 1L),
    "Hypertension (treated)"          = pct(df$treated_htn),
    "Atrial fibrillation"             = pct(df$af),
    "CKD stage 3–5"                   = pct(df$ckd),
    "Heart failure"                   = pct(df$hf),
    "IHD"                             = vt("IHD"),
    "Stroke"                          = vt("Stroke"),
    "PAD"                             = vt("PAD"),
    "LDL candidate"                   = pct(df$ldl_cand),
    "SBP candidate"                   = pct(df$sbp_cand)
  )
}

all_cand_tbl <- dplyr::bind_rows(dm_cand_tbl, ascvd_cand_tbl)

ov_u  <- summ_cand(dplyr::filter(all_cand_tbl,  tx_status == "Untreated"),    is_dm = FALSE)
ov_ut <- summ_cand(dplyr::filter(all_cand_tbl,  tx_status == "Undertreated"), is_dm = FALSE)
dm_u  <- summ_cand(dplyr::filter(dm_cand_tbl,   tx_status == "Untreated"),    is_dm = TRUE)
dm_ut <- summ_cand(dplyr::filter(dm_cand_tbl,   tx_status == "Undertreated"), is_dm = TRUE)
ax_u  <- summ_cand(dplyr::filter(ascvd_cand_tbl, tx_status == "Untreated"),   is_dm = FALSE)
ax_ut <- summ_cand(dplyr::filter(ascvd_cand_tbl, tx_status == "Undertreated"), is_dm = FALSE)

cand_tbl <- data.frame(
  Characteristic  = names(ov_u),
  OV_Untreated    = unlist(ov_u),
  OV_Undertreated = unlist(ov_ut),
  DM_Untreated    = unlist(dm_u),
  DM_Undertreated = unlist(dm_ut),
  AX_Untreated    = unlist(ax_u),
  AX_Undertreated = unlist(ax_ut),
  stringsAsFactors = FALSE, row.names = NULL
)

kbl(cand_tbl, align = c("l","c","c","c","c","c","c"),
    col.names = c("Characteristic",
                  "Untreated", "Undertreated",
                  "Untreated", "Undertreated",
                  "Untreated", "Undertreated")) |>
  kable_styling(
    bootstrap_options = c("striped","hover","condensed"),
    full_width = FALSE, position = "left"
  ) |>
  add_header_above(c(" " = 1,
                     "Overall" = 2,
                     "DM (primary)" = 2,
                     "ASCVD (secondary)" = 2)) |>
  column_spec(1, width = "16em") |>
  column_spec(2:7, width = "7em")
```

**Supplementary Table 1.** Baseline characteristics of actionable treatment candidates by treatment status.

<br>
<br>

```{r suppl_table2, echo=FALSE}
wimd_lvls  <- 1:5
wimd_hdrs  <- c("1 (most deprived)", "2", "3", "4", "5 (least deprived)")

# Build one summary column per cohort × WIMD quintile
ax_wimd <- lapply(wimd_lvls, function(w)
  summ_cand(dplyr::filter(ascvd_cand_tbl, wimd == w), is_dm = FALSE))
dm_wimd <- lapply(wimd_lvls, function(w)
  summ_cand(dplyr::filter(dm_cand_tbl, wimd == w), is_dm = TRUE))

wimd_tbl <- data.frame(
  Characteristic = names(ax_wimd[[1]]),
  setNames(lapply(ax_wimd, unlist), paste0("AX_W", wimd_lvls)),
  setNames(lapply(dm_wimd, unlist), paste0("DM_W", wimd_lvls)),
  stringsAsFactors = FALSE, row.names = NULL
)

kbl(wimd_tbl,
    align  = c("l", rep("c", 10)),
    col.names = c("Characteristic", wimd_hdrs, wimd_hdrs)) |>
  kable_styling(
    bootstrap_options = c("striped","hover","condensed"),
    full_width = FALSE, position = "left"
  ) |>
  add_header_above(c(" " = 1,
                     "ASCVD (secondary)" = 5,
                     "DM (primary)"      = 5)) |>
  column_spec(1, width = "15em") |>
  column_spec(2:11, width = "6em")
```

**Supplementary Table 2.** Baseline characteristics of actionable treatment candidates by WIMD quintile. 

<br>
<br>

```{r suppl_table3, echo=FALSE}
rq_lvls <- c("Q1 (lowest)", "Q2", "Q3", "Q4 (highest)")
rq_hdrs <- c("Q1 (lowest)", "Q2", "Q3", "Q4 (highest)")

ax_rq <- lapply(rq_lvls, function(q)
  summ_cand(dplyr::filter(ascvd_cand_tbl, risk_q == q), is_dm = FALSE))
dm_rq <- lapply(rq_lvls, function(q)
  summ_cand(dplyr::filter(dm_cand_tbl,   risk_q == q), is_dm = TRUE))

rq_tbl <- data.frame(
  Characteristic = names(ax_rq[[1]]),
  setNames(lapply(ax_rq, unlist), paste0("AX_", rq_lvls)),
  setNames(lapply(dm_rq,  unlist), paste0("DM_", rq_lvls)),
  stringsAsFactors = FALSE, row.names = NULL
)

kbl(rq_tbl,
    align  = c("l", rep("c", 8)),
    col.names = c("Characteristic", rq_hdrs, rq_hdrs)) |>
  kable_styling(
    bootstrap_options = c("striped","hover","condensed"),
    full_width = FALSE, position = "left"
  ) |>
  add_header_above(c(" " = 1,
                     "ASCVD (secondary)" = 4,
                     "DM (primary)"      = 4)) |>
  column_spec(1, width = "15em") |>
  column_spec(2:9, width = "7em")
```

**Supplementary Table 3.** Baseline characteristics of actionable treatment candidates by baseline cardiovascular risk quartile.

<br>
<br>

```{r fig_wimd, include=FALSE}
wimd_labels <- c("1 (most\ndeprived)", "2", "3", "4", "5 (least\ndeprived)")
wimd_pal    <- setNames(
  c("#D62728", "#FF7F0E", "#9467BD", "#17BECF", "#2CA02C"),
  wimd_labels
)

dm_wimd    <- dm_scored    |>
  dplyr::filter(!is.na(wimd), wimd %in% 1:5) |>
  dplyr::mutate(wimd_f = factor(wimd, levels = 1:5, labels = wimd_labels))
ascvd_wimd <- ascvd_scored |>
  dplyr::filter(!is.na(wimd), wimd %in% 1:5) |>
  dplyr::mutate(wimd_f = factor(wimd, levels = 1:5, labels = wimd_labels))

# ── Summary statistics by WIMD ─────────────────────────────────────────────────
wimd_sum <- dplyr::bind_rows(
  dm_wimd |>
    dplyr::group_by(wimd_f) |>
    dplyr::summarise(
      pct_ldl_target = 100 * mean(ldl <= 2.0,             na.rm = TRUE),
      pct_sbp_target = 100 * mean(sbp <= 130,             na.rm = TRUE),
      pct_on_statin  = 100 * mean(statin_rx_last6m == 1L, na.rm = TRUE),
      pct_on_aht     = 100 * mean(treated_htn == 1L,      na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::mutate(Cohort = "DM (primary)"),
  ascvd_wimd |>
    dplyr::group_by(wimd_f) |>
    dplyr::summarise(
      pct_ldl_target = 100 * mean(ldl <= 1.8,             na.rm = TRUE),
      pct_sbp_target = 100 * mean(sbp <= 130,             na.rm = TRUE),
      pct_on_statin  = 100 * mean(statin_rx_last6m == 1L, na.rm = TRUE),
      pct_on_aht     = 100 * mean(treated_htn == 1L,      na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::mutate(Cohort = "ASCVD (secondary)")
)

# ── Panel A: QRISK2 by WIMD (DM) ──────────────────────────────────────────────
pWA <- ggplot(dm_wimd |> dplyr::filter(!is.na(qrisk10y)),
              aes(x = qrisk10y, colour = wimd_f, fill = wimd_f)) +
  geom_density(alpha = 0.18, linewidth = 0.65) +
  scale_colour_manual(values = wimd_pal, name = "WIMD quintile") +
  scale_fill_manual(  values = wimd_pal, name = "WIMD quintile") +
  labs(tag = "A", title= "DM (primary)",
       x = "10-year risk (%)", y = "Density") +
  theme_cap() +
  theme(legend.position = "none", plot.tag = element_text(face = "bold", size = 13))

# ── Panel B: REACT by WIMD (ASCVD) ────────────────────────────────────────────
pWB <- ggplot(ascvd_wimd |> dplyr::filter(!is.na(react5y)),
              aes(x = react5y, colour = wimd_f, fill = wimd_f)) +
  geom_density(alpha = 0.18, linewidth = 0.65) +
  scale_colour_manual(values = wimd_pal, name = "WIMD quintile") +
  scale_fill_manual(  values = wimd_pal, name = "WIMD quintile") +
  labs(tag = "B", title = "ASCVD (secondary)",
       x = "5-year risk (%)", y = "Density") +
  theme_cap() +
  theme(axis.title.y = element_blank(), legend.position = "right", plot.tag = element_text(face = "bold", size = 13))

# ── Panels C/D: all 4 metrics by WIMD, split by cohort ────────────────────────
wimd_all_long <- wimd_sum |>
  tidyr::pivot_longer(
    cols     = c(pct_ldl_target, pct_on_statin, pct_sbp_target, pct_on_aht),
    names_to = "Metric", values_to = "Pct"
  ) |>
  dplyr::mutate(
    Metric = factor(Metric,
                    levels = c("pct_ldl_target", "pct_on_statin",
                               "pct_sbp_target", "pct_on_aht"),
                    labels = c("At LDL target", "On LLT",
                               "At SBP target", "On AHT"))
  )

.metric_lbl_ldl <- "At LDL target\n(<2.0 / <1.8 mmol/L)"
.metric_lbl_sbp <- "At SBP target\n(≤130 mmHg)"
.metric_pal <- c(setNames("#CB181D", .metric_lbl_ldl), "On LLT" = "#FC9272",
                 setNames("#2171B5", .metric_lbl_sbp), "On AHT" = "#9ECAE1")
.metric_shp <- c(setNames(16, .metric_lbl_ldl), "On LLT" = 17,
                 setNames(15, .metric_lbl_sbp), "On AHT" = 18)
.metric_lvls <- c(.metric_lbl_ldl, "On LLT", .metric_lbl_sbp, "On AHT")

.recode_metric <- function(dat) {
  dat |> dplyr::mutate(
    Metric = factor(
      dplyr::recode(as.character(Metric),
                    "At LDL target" = .metric_lbl_ldl,
                    "At SBP target" = .metric_lbl_sbp),
      levels = .metric_lvls
    )
  )
}

.make_wimd_all <- function(dat, tag, show_legend = FALSE) {
  dat <- .recode_metric(dat)
  p <- ggplot(dat, aes(x = wimd_f, y = Pct,
                       colour = Metric, group = Metric, shape = Metric)) +
    geom_line(linewidth = 1.0) +
    geom_point(size = 3.2) +
    scale_colour_manual(values = .metric_pal, name = "Metric") +
    scale_shape_manual(values  = .metric_shp, name = "Metric") +
    scale_y_continuous(labels = scales::label_percent(scale = 1),
                       limits = c(0, 100),
                       expand = expansion(mult = c(0.02, 0.14))) +
    labs(tag = tag, x = "WIMD quintile", y = "Proportion") +
    guides(colour = guide_legend(nrow = 4), shape = guide_legend(nrow = 4)) +
    theme_cap() +
    theme(legend.text       = element_text(size = 8),
          legend.key.width  = unit(1.8, "lines"),
          legend.key.height = unit(2.2, "lines"),
          plot.tag          = element_text(face = "bold", size = 13))
  if (show_legend) p + theme(legend.position = "right")
  else             p + theme(legend.position = "none")
}

pWC <- .make_wimd_all(
  wimd_all_long |> dplyr::filter(Cohort == "DM (primary)"),   "C"
)
pWD <- .make_wimd_all(
  wimd_all_long |> dplyr::filter(Cohort == "ASCVD (secondary)"), "D", show_legend = TRUE
) + theme(axis.title.y = element_blank())

fig5 <- (pWA | pWB) / (pWC | pWD) +
  plot_layout(heights = c(1.2, 1))
fig5
ggsave("figures/fig5.tiff", fig5, width = 11, height = 11 * 0.75, dpi = 300, compression = "lzw")
```

<br>
<br>

```{r suppl_fig1, fig.width=11, fig.asp=0.75, echo=FALSE}
fig5
```

**Supplementary Figure 1.** Estimated cardiovascular risk and risk factor control by WIMD quintile. (A) QRISK2 risk, DM cohort. (B) REACT risk, ASCVD cohort. (C–D) LDL-C and SBP target attainment and treatment receipt.

<br>
<br>

```{r fig_riskq, include=FALSE}
q_pal <- c("#2171B5", "#6BAED6", "#F16913", "#CB181D")

q_dm_breaks    <- quantile(dm_scored$qrisk10y,  probs = c(0, .25, .5, .75, 1), na.rm = TRUE)
q_ascvd_breaks <- quantile(ascvd_scored$react5y, probs = c(0, .25, .5, .75, 1), na.rm = TRUE)
q_labs         <- c("Q1 (lowest)", "Q2", "Q3", "Q4 (highest)")
q_labs_dm      <- .q_range_labels(q_dm_breaks)
q_labs_ascvd   <- .q_range_labels(q_ascvd_breaks)

dm_q    <- dm_scored |>
  dplyr::filter(!is.na(qrisk10y)) |>
  dplyr::mutate(risk_q = cut(qrisk10y,  breaks = q_dm_breaks,
                              labels = q_labs_dm, include.lowest = TRUE))
ascvd_q <- ascvd_scored |>
  dplyr::filter(!is.na(react5y)) |>
  dplyr::mutate(risk_q = cut(react5y, breaks = q_ascvd_breaks,
                               labels = q_labs_ascvd, include.lowest = TRUE))

# ── Quartile-level summaries ───────────────────────────────────────────────────
rq_sum <- dplyr::bind_rows(
  dm_q |>
    dplyr::group_by(risk_q) |>
    dplyr::summarise(
      pct_ldl_target = 100 * mean(ldl <= 2.0,             na.rm = TRUE),
      pct_sbp_target = 100 * mean(sbp <= 130,             na.rm = TRUE),
      pct_on_statin  = 100 * mean(statin_rx_last6m == 1L, na.rm = TRUE),
      pct_on_aht     = 100 * mean(treated_htn == 1L,      na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::mutate(Cohort = "DM (primary)"),
  ascvd_q |>
    dplyr::group_by(risk_q) |>
    dplyr::summarise(
      pct_ldl_target = 100 * mean(ldl <= 1.8,             na.rm = TRUE),
      pct_sbp_target = 100 * mean(sbp <= 130,             na.rm = TRUE),
      pct_on_statin  = 100 * mean(statin_rx_last6m == 1L, na.rm = TRUE),
      pct_on_aht     = 100 * mean(treated_htn == 1L,      na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::mutate(Cohort = "ASCVD (secondary)")
) |>
  dplyr::filter(!is.na(risk_q))

# ── Panel A: LDL-C by risk quartile (DM) ──────────────────────────────────────
pQA <- ggplot(dm_q |> dplyr::filter(!is.na(ldl), !is.na(risk_q)),
              aes(x = ldl, colour = risk_q, fill = risk_q)) +
  geom_density(alpha = 0.20, linewidth = 0.7) +
  geom_vline(xintercept = 2.0, linetype = "dashed", colour = "#555", linewidth = 0.6) +
  annotate("text", x = 2.07, y = Inf, vjust = 1.8, hjust = 0,
           label = "2.0 mmol/L", size = 2.4, colour = "#555") +
  scale_colour_manual(values = q_pal, name = "Risk quartile") +
  scale_fill_manual(  values = q_pal, name = "Risk quartile") +
  labs(tag = "A", title = "DM (primary)",
       x = "LDL-C (mmol/L)", y = "Density") +
  theme_cap() +
  theme(legend.position = "none", plot.tag = element_text(face = "bold", size = 13))

# ── Panel B: LDL-C by risk quartile (ASCVD) ───────────────────────────────────
pQB <- ggplot(ascvd_q |> dplyr::filter(!is.na(ldl), !is.na(risk_q)),
              aes(x = ldl, colour = risk_q, fill = risk_q)) +
  geom_density(alpha = 0.20, linewidth = 0.7) +
  geom_vline(xintercept = 1.8, linetype = "dashed", colour = "#555", linewidth = 0.6) +
  annotate("text", x = 1.87, y = Inf, vjust = 1.8, hjust = 0,
           label = "1.8 mmol/L", size = 2.4, colour = "#555") +
  scale_colour_manual(values = q_pal, name = "Risk quartile") +
  scale_fill_manual(  values = q_pal, name = "Risk quartile") +
  labs(tag = "B", title = "ASCVD (secondary)",
       x = "LDL-C (mmol/L)", y = "Density") +
  theme_cap() +
  theme(axis.title.y = element_blank(), legend.position = "right", legend.key.height = unit(2.2, "lines"),
        plot.tag = element_text(face = "bold", size = 13))

# ── Panel C: SBP by risk quartile (DM) ────────────────────────────────────────
pQC <- ggplot(dm_q |> dplyr::filter(!is.na(sbp), !is.na(risk_q)),
              aes(x = sbp, colour = risk_q, fill = risk_q)) +
  geom_density(alpha = 0.20, linewidth = 0.7) +
  geom_vline(xintercept = 130, linetype = "dashed", colour = "#555", linewidth = 0.6) +
  annotate("text", x = 131, y = Inf, vjust = 1.8, hjust = 0,
           label = "≤130 mmHg", size = 2.4, colour = "#555") +
  scale_colour_manual(values = q_pal, name = "Risk quartile") +
  scale_fill_manual(  values = q_pal, name = "Risk quartile") +
  labs(tag = "C",
       x = "SBP (mmHg)", y = "Density") +
  theme_cap() +
  theme(legend.position = "none", plot.tag = element_text(face = "bold", size = 13))

# ── Panel D: SBP by risk quartile (ASCVD) ─────────────────────────────────────
pQD <- ggplot(ascvd_q |> dplyr::filter(!is.na(sbp), !is.na(risk_q)),
              aes(x = sbp, colour = risk_q, fill = risk_q)) +
  geom_density(alpha = 0.20, linewidth = 0.7) +
  geom_vline(xintercept = 130, linetype = "dashed", colour = "#555", linewidth = 0.6) +
  annotate("text", x = 131, y = Inf, vjust = 1.8, hjust = 0,
           label = "≤130 mmHg", size = 2.4, colour = "#555") +
  scale_colour_manual(values = q_pal, name = "Risk quartile") +
  scale_fill_manual(  values = q_pal, name = "Risk quartile") +
  labs(tag = "D", 
       x = "SBP (mmHg)", y = "Density") +
  theme_cap() +
  theme(axis.title.y = element_blank(), legend.position = "right", legend.key.height = unit(2.2, "lines"),
        plot.tag = element_text(face = "bold", size = 13))

# ── Panels E/F: Treatment attainment and rates by risk quartile, split by cohort
rq_long <- rq_sum |>
  tidyr::pivot_longer(
    cols     = c(pct_ldl_target, pct_sbp_target, pct_on_statin, pct_on_aht),
    names_to = "Metric", values_to = "Pct"
  ) |>
  dplyr::mutate(
    Metric = factor(Metric,
                    levels = c("pct_ldl_target", "pct_on_statin",
                               "pct_sbp_target", "pct_on_aht"),
                    labels = c("At LDL target", "On LLT",
                               "At SBP target", "On AHT"))
  )

.make_rq_panel <- function(dat, tag, subtitle, show_legend = FALSE) {
  dat <- dat |>
    dplyr::filter(!is.na(risk_q)) |>
    dplyr::mutate(risk_q = droplevels(risk_q)) |>
    .recode_metric()
  p <- ggplot(dat, aes(x = risk_q, y = Pct,
               colour = Metric, group = Metric, shape = Metric)) +
    geom_line(linewidth = 1.0) +
    geom_point(size = 3.0) +
    scale_colour_manual(values = .metric_pal, name = "Metric") +
    scale_shape_manual(values  = .metric_shp, name = "Metric") +
    scale_y_continuous(labels = scales::label_percent(scale = 1),
                       limits = c(0, 100),
                       expand = expansion(mult = c(0.02, 0.12))) +
    labs(tag = tag,
         x = "Baseline risk quartile", y = "Proportion") +
    guides(colour = guide_legend(nrow = 4), shape = guide_legend(nrow = 4)) +
    theme_cap() +
    theme(legend.text       = element_text(size = 8),
          legend.key.width  = unit(2, "lines"),
          legend.key.height = unit(2.2, "lines"),
          plot.tag          = element_text(face = "bold", size = 13))
  if (show_legend) p + theme(legend.position = "right")
  else             p + theme(legend.position = "none")
}

pQE <- .make_rq_panel(
  rq_long |> dplyr::filter(Cohort == "DM (primary)"),
  "E", "DM (primary)"
)
pQF <- .make_rq_panel(
  rq_long |> dplyr::filter(Cohort == "ASCVD (secondary)"),
  "F", "ASCVD (secondary)", show_legend = TRUE
) + theme(axis.title.y = element_blank())

fig6 <- (pQA | pQB) / (pQC | pQD) / (pQE | pQF) +
  plot_layout(heights = c(1, 1, 0.9))
fig6
ggsave("figures/fig6.tiff", fig6, width = 11, height = 11 * 1.10, dpi = 300, compression = "lzw")
```

<br>
<br>

```{r suppl_fig2, fig.width=11, fig.asp=1.10, echo=FALSE}
fig6
```

**Supplementary Figure 2.** Risk factor control and treatment intensity by baseline risk quartile. (A–B) LDL-C distributions. (C–D) SBP distributions. (E–F) LDL-C and SBP target attainment and treatment receipt.

<br>
<br>

```{r sex_data, echo=FALSE}
sex_pal <- c("Male" = "#2171B5", "Female" = "#CB181D")

dm_sex    <- dm_scored    |>
  dplyr::filter(!is.na(sex)) |>
  dplyr::mutate(sex_f = factor(dplyr::if_else(sex == "M", "Male", "Female"),
                               levels = c("Male", "Female")))
ascvd_sex <- ascvd_scored |>
  dplyr::filter(!is.na(sex)) |>
  dplyr::mutate(sex_f = factor(dplyr::if_else(sex == "M", "Male", "Female"),
                               levels = c("Male", "Female")))

sex_sum <- dplyr::bind_rows(
  dm_sex |>
    dplyr::group_by(sex_f) |>
    dplyr::summarise(
      pct_ldl_target = 100 * mean(ldl <= 2.0,             na.rm = TRUE),
      pct_sbp_target = 100 * mean(sbp <= 130,             na.rm = TRUE),
      pct_on_statin  = 100 * mean(statin_rx_last6m == 1L, na.rm = TRUE),
      pct_on_aht     = 100 * mean(treated_htn == 1L,      na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::mutate(Cohort = "DM (primary)"),
  ascvd_sex |>
    dplyr::group_by(sex_f) |>
    dplyr::summarise(
      pct_ldl_target = 100 * mean(ldl <= 1.8,             na.rm = TRUE),
      pct_sbp_target = 100 * mean(sbp <= 130,             na.rm = TRUE),
      pct_on_statin  = 100 * mean(statin_rx_last6m == 1L, na.rm = TRUE),
      pct_on_aht     = 100 * mean(treated_htn == 1L,      na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::mutate(Cohort = "ASCVD (secondary)")
)

.sv <- function(co, sx, mt)
  round(sex_sum[[mt]][sex_sum$Cohort == co & sex_sum$sex_f == sx], 1)
pct_statin_f_dm  <- .sv("DM (primary)",      "Female", "pct_on_statin")
pct_statin_m_dm  <- .sv("DM (primary)",      "Male",   "pct_on_statin")
pct_statin_f_asc <- .sv("ASCVD (secondary)", "Female", "pct_on_statin")
pct_statin_m_asc <- .sv("ASCVD (secondary)", "Male",   "pct_on_statin")
pct_ldl_f_dm     <- .sv("DM (primary)",      "Female", "pct_ldl_target")
pct_ldl_m_dm     <- .sv("DM (primary)",      "Male",   "pct_ldl_target")
pct_ldl_f_asc    <- .sv("ASCVD (secondary)", "Female", "pct_ldl_target")
pct_ldl_m_asc    <- .sv("ASCVD (secondary)", "Male",   "pct_ldl_target")
```

```{r fig_sex, include=FALSE}
# ── Panels A/B: LDL-C distributions by sex ────────────────────────────────
pSA <- ggplot(dm_sex |> dplyr::filter(!is.na(ldl)),
              aes(x = ldl, colour = sex_f, fill = sex_f)) +
  geom_density(alpha = 0.25, linewidth = 0.7) +
  geom_vline(xintercept = 2.0, linetype = "dashed", colour = "#555", linewidth = 0.6) +
  annotate("text", x = 2.07, y = Inf, vjust = 1.8, hjust = 0,
           label = "2.0 mmol/L", size = 2.4, colour = "#555") +
  scale_colour_manual(values = sex_pal, name = "Sex") +
  scale_fill_manual(  values = sex_pal, name = "Sex") +
  labs(tag = "A", title = "DM (primary)", x = "LDL-C (mmol/L)", y = "Density") +
  theme_cap() +
  theme(legend.position = "none", plot.tag = element_text(face = "bold", size = 13))

pSB <- ggplot(ascvd_sex |> dplyr::filter(!is.na(ldl)),
              aes(x = ldl, colour = sex_f, fill = sex_f)) +
  geom_density(alpha = 0.25, linewidth = 0.7) +
  geom_vline(xintercept = 1.8, linetype = "dashed", colour = "#555", linewidth = 0.6) +
  annotate("text", x = 1.87, y = Inf, vjust = 1.8, hjust = 0,
           label = "1.8 mmol/L", size = 2.4, colour = "#555") +
  scale_colour_manual(values = sex_pal, name = "Sex") +
  scale_fill_manual(  values = sex_pal, name = "Sex") +
  labs(tag = "B", title = "ASCVD (secondary)", x = "LDL-C (mmol/L)", y = "Density") +
  theme_cap() +
  theme(legend.position = "right", axis.title.y = element_blank(),
        plot.tag = element_text(face = "bold", size = 13))

# ── Panels C/D: SBP distributions by sex ──────────────────────────────────
pSC <- ggplot(dm_sex |> dplyr::filter(!is.na(sbp)),
              aes(x = sbp, colour = sex_f, fill = sex_f)) +
  geom_density(alpha = 0.25, linewidth = 0.7) +
  geom_vline(xintercept = 130, linetype = "dashed", colour = "#555", linewidth = 0.6) +
  annotate("text", x = 131, y = Inf, vjust = 1.8, hjust = 0,
           label = "≤130 mmHg", size = 2.4, colour = "#555") +
  scale_colour_manual(values = sex_pal, name = "Sex") +
  scale_fill_manual(  values = sex_pal, name = "Sex") +
  labs(tag = "C", x = "SBP (mmHg)", y = "Density") +
  theme_cap() +
  theme(legend.position = "none", plot.tag = element_text(face = "bold", size = 13))

pSD <- ggplot(ascvd_sex |> dplyr::filter(!is.na(sbp)),
              aes(x = sbp, colour = sex_f, fill = sex_f)) +
  geom_density(alpha = 0.25, linewidth = 0.7) +
  geom_vline(xintercept = 130, linetype = "dashed", colour = "#555", linewidth = 0.6) +
  annotate("text", x = 131, y = Inf, vjust = 1.8, hjust = 0,
           label = "≤130 mmHg", size = 2.4, colour = "#555") +
  scale_colour_manual(values = sex_pal, name = "Sex") +
  scale_fill_manual(  values = sex_pal, name = "Sex") +
  labs(tag = "D", x = "SBP (mmHg)", y = "Density") +
  theme_cap() +
  theme(legend.position = "right", axis.title.y = element_blank(),
        plot.tag = element_text(face = "bold", size = 13))

# ── Panels E/F: treatment metrics by sex (dodged bars) ────────────────────
sex_long <- sex_sum |>
  tidyr::pivot_longer(
    cols     = c(pct_ldl_target, pct_on_statin, pct_sbp_target, pct_on_aht),
    names_to = "Metric", values_to = "Pct"
  ) |>
  dplyr::mutate(
    Metric = factor(Metric,
                    levels = c("pct_ldl_target", "pct_on_statin",
                               "pct_sbp_target", "pct_on_aht"),
                    labels = c("At LDL target", "On LLT",
                               "At SBP target", "On AHT"))
  ) |>
  .recode_metric()

.sex_bar <- function(cohort_nm, tag, show_legend = FALSE) {
  d <- sex_long |> dplyr::filter(Cohort == cohort_nm)
  p <- ggplot(d, aes(x = Metric, y = Pct, fill = sex_f)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.65) +
    scale_fill_manual(values = sex_pal, name = "Sex") +
    scale_y_continuous(labels = scales::label_percent(scale = 1),
                       limits = c(0, 100),
                       expand = expansion(mult = c(0, 0.08))) +
    labs(tag = tag, x = NULL, y = "Proportion") +
    theme_cap() +
    theme(axis.text.x = element_text(size = 8),
          plot.tag    = element_text(face = "bold", size = 13))
  if (show_legend) p + theme(legend.position = "right")
  else             p + theme(legend.position = "none")
}

pSE <- .sex_bar("DM (primary)",      "E")
pSF <- .sex_bar("ASCVD (secondary)", "F", show_legend = TRUE)

fig7 <- (pSA | pSB) / (pSC | pSD) / (pSE | pSF) +
  plot_layout(heights = c(1, 1, 0.9))
fig7
ggsave("figures/fig7.tiff", fig7, width = 11, height = 11 * 1.1, dpi = 300, compression = "lzw")
```

<br>
<br>

```{r suppl_fig3, fig.width=11, fig.asp=1.1, echo=FALSE}
fig7
```

```{r, echo=FALSE}

ggsave(filename = 'fig7.tiff', plot = fig7, device = 'tiff', width=11, height=12, units ='in', dpi=300, compression = 'lzw')

```

**Supplementary Figure 3.** Risk factor control and treatment intensity by baseline risk quartile by sex. **(A, B)** LDL-C distributions; dashed lines mark guideline targets. **(C, D)** SBP distributions; dashed line marks 130 mmHg. **(E, F)** Target attainment and treatment receipt.

<br>
<br>

## Base tables

```{r rt_setup, include=FALSE}
.mask_n <- function(x) dplyr::if_else(x < 10L, 10L, x)
.mask_all_n <- function(df) dplyr::mutate(df, dplyr::across(
  dplyr::matches("^N$|^N_|^N "), .mask_n))
```

<br>

### RT1. Figure 1 panels A–B — Cardiovascular risk score distributions
```{r rt1_risk, echo=FALSE}
rt1 <- dplyr::bind_rows(
  dm_scored |> dplyr::filter(!is.na(qrisk10y)) |>
    dplyr::summarise(Cohort = "DM (primary)", Metric = "QRISK2 10y (%)",
      N = dplyr::n(), Mean = round(mean(qrisk10y), 1), SD = round(sd(qrisk10y), 1),
      Median = round(median(qrisk10y), 1),
      Q1 = round(quantile(qrisk10y, 0.25), 1), Q3 = round(quantile(qrisk10y, 0.75), 1)),
  ascvd_scored |> dplyr::filter(!is.na(react5y)) |>
    dplyr::summarise(Cohort = "ASCVD (secondary)", Metric = "REACT 5y (%)",
      N = dplyr::n(), Mean = round(mean(react5y), 1), SD = round(sd(react5y), 1),
      Median = round(median(react5y), 1),
      Q1 = round(quantile(react5y, 0.25), 1), Q3 = round(quantile(react5y, 0.75), 1))
) |> .mask_all_n()
kbl(rt1) |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```


<br>

### RT2. Figure 1 panels C–D — LDL-C and SBP distributions
```{r rt2_biomarkers, echo=FALSE}
.rt_summ <- function(df, cohort, metric, col) {
  df |> dplyr::filter(!is.na(.data[[col]])) |>
    dplyr::summarise(
      N      = dplyr::n(),
      Mean   = round(mean(.data[[col]]), 2),
      SD     = round(sd(.data[[col]]), 2),
      Median = round(median(.data[[col]]), 2),
      Q1     = round(quantile(.data[[col]], 0.25), 2),
      Q3     = round(quantile(.data[[col]], 0.75), 2)
    ) |>
    dplyr::mutate(Cohort = cohort, Metric = metric, .before = 1)
}
rt2 <- dplyr::bind_rows(
  .rt_summ(dm_scored,    "DM (primary)",      "LDL-C (mmol/L)", "ldl"),
  .rt_summ(ascvd_scored, "ASCVD (secondary)", "LDL-C (mmol/L)", "ldl"),
  .rt_summ(dm_scored,    "DM (primary)",      "SBP (mmHg)",     "sbp"),
  .rt_summ(ascvd_scored, "ASCVD (secondary)", "SBP (mmHg)",     "sbp")
) |> .mask_all_n()
kbl(rt2) |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```


<br>

### RT3. Figure 1 panels E–F — Treatment intensity proportions
```{r rt3_treatment, echo=FALSE}
rt3 <- treatment_rates |>
  dplyr::mutate(
    N_cohort   = dplyr::if_else(Cohort == "DM (primary)", nrow(dm_base), nrow(ascvd_base)),
    N_category = round(N_cohort * Pct / 100)
  ) |>
  dplyr::select(Cohort, Treatment, Category = Label,
                `N (cohort)` = N_cohort, `N (category)` = N_category, `%` = Pct) |>
  dplyr::mutate(`%` = round(`%`, 1)) |>
  .mask_all_n()
kbl(rt3) |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```


<br>

### RT4. Figure 2 panels A–B — Risk distributions under modelled scenarios
```{r rt4_scenarios, echo=FALSE}
.n_valid <- function(df, col) nrow(dplyr::filter(df, !is.na(.data[[col]])))
rt4 <- dplyr::bind_rows(
  scen_means_dm    |> dplyr::mutate(Cohort = "DM (primary)",      Risk = "QRISK2 10y (%)"),
  scen_means_ascvd |> dplyr::mutate(Cohort = "ASCVD (secondary)", Risk = "REACT 5y (%)")
) |>
  dplyr::mutate(N = dplyr::if_else(Cohort == "DM (primary)",
                                    .n_valid(dm_scored,    "qrisk10y"),
                                    .n_valid(ascvd_scored, "react5y"))) |>
  dplyr::select(Cohort, Risk, Scenario = scenario, N, Mean = mean_risk, Median = median_risk) |>
  dplyr::mutate(dplyr::across(c(Mean, Median), \(x) round(x, 1))) |>
  .mask_all_n()
kbl(rt4) |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```


<br>

### RT5. Figure 2 panels C–D — Events expected per 1,000 by risk quartile and scenario
```{r rt5_events_expected, echo=FALSE}
.n_by_q <- function(df, q_col) {
  df |> dplyr::filter(!is.na(.data[[q_col]])) |>
    dplyr::count(.data[[q_col]], name = "N") |>
    dplyr::rename(risk_q = 1)
}
n_q_dm5    <- .n_by_q(dm_opp3,    "risk_q") |> dplyr::mutate(Cohort = "DM (primary)")
n_q_ascvd5 <- .n_by_q(ascvd_opp3, "risk_q") |> dplyr::mutate(Cohort = "ASCVD (secondary)")

rt5 <- opp3_central |>
  dplyr::left_join(dplyr::bind_rows(n_q_dm5, n_q_ascvd5), by = c("Cohort","risk_q")) |>
  dplyr::select(Cohort, Risk_quartile = risk_q, N, Scenario = scenario,
                Events_expected_per_1000 = events_expected) |>
  dplyr::mutate(Events_expected_per_1000 = round(Events_expected_per_1000, 1)) |>
  dplyr::arrange(Cohort, Risk_quartile, Scenario) |>
  .mask_all_n()
kbl(rt5) |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```


<br>

### RT6. Figure 2 panels E–F — Events prevented per 1,000 by risk quartile (with 95% CI)
```{r rt6_events_prevented, echo=FALSE}
rt6 <- dplyr::left_join(
  opp3_stacked |>
    dplyr::select(Cohort, risk_q, Component = component,
                  Events_prevented_per_1000 = events),
  opp3_bar_err |>
    dplyr::select(Cohort, risk_q, CI_lower = err_lo, CI_upper = err_hi),
  by = c("Cohort","risk_q")
) |>
  dplyr::left_join(dplyr::bind_rows(n_q_dm5, n_q_ascvd5), by = c("Cohort","risk_q")) |>
  dplyr::select(Cohort, Risk_quartile = risk_q, N, Component,
                Events_prevented_per_1000, CI_lower, CI_upper) |>
  dplyr::mutate(dplyr::across(c(Events_prevented_per_1000, CI_lower, CI_upper),
                               \(x) round(x, 1))) |>
  dplyr::arrange(Cohort, Risk_quartile, Component) |>
  .mask_all_n()
kbl(rt6) |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```


<br>

### RT7. Figure 3 panels A–B — LDL-C treatment status by risk quartile
```{r rt7_ldl_status, echo=FALSE}
rt7 <- dplyr::bind_rows(ldl_status_dm, ldl_status_ascvd) |>
  dplyr::select(Cohort, Risk_quartile = risk_q_lbl, Status = ldl_status, N = n, `%` = pct) |>
  dplyr::mutate(`%` = round(`%`, 1)) |>
  dplyr::arrange(Cohort, Risk_quartile, Status) |>
  .mask_all_n()
kbl(rt7) |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```


<br>

### RT8. Figure 3 panels C–D — SBP treatment status by risk quartile
```{r rt8_sbp_status, echo=FALSE}
rt8 <- dplyr::bind_rows(sbp_status_dm, sbp_status_ascvd) |>
  dplyr::select(Cohort, Risk_quartile = risk_q_lbl, Status = sbp_status, N = n, `%` = pct) |>
  dplyr::mutate(`%` = round(`%`, 1)) |>
  dplyr::arrange(Cohort, Risk_quartile, Status) |>
  .mask_all_n()
kbl(rt8) |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```


<br>

### RT9. Figure 3 panels E–F — Events prevented per 1,000 candidates by risk quartile (with 95% CI)
```{r rt9_cand_ep, echo=FALSE}
rt9_totals <- dplyr::bind_rows(cand_ep_dm, cand_ep_ascvd) |>
  dplyr::select(Cohort, Risk_quartile = risk_q_lbl, N_candidates = n_cand,
                Events_prevented_per_1000 = ep_per1k, CI_lower = err_lo, CI_upper = err_hi) |>
  dplyr::mutate(dplyr::across(c(Events_prevented_per_1000, CI_lower, CI_upper),
                               \(x) round(x, 1))) |>
  .mask_all_n()
rt9_components <- dplyr::bind_rows(cand_stacked_dm, cand_stacked_ascvd) |>
  dplyr::select(Cohort, Risk_quartile = risk_q_lbl, Component = component,
                Events_prevented_per_1000 = ep_per1k) |>
  dplyr::mutate(Events_prevented_per_1000 = round(Events_prevented_per_1000, 1))
kbl(rt9_totals,
    caption = "Total events prevented per 1,000 candidates with 95% CI (error bar limits in figure)") |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
kbl(rt9_components,
    caption = "Stacked bar components: all-candidates-treated vs additional-to-target") |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```


<br>

### RT10. Supplementary Figure 1 — Treatment receipt by WIMD quintile (with N per quintile)
```{r rt10_wimd, echo=FALSE}
n_wimd <- dplyr::bind_rows(
  dm_wimd    |> dplyr::count(wimd_f, name = "N") |> dplyr::mutate(Cohort = "DM (primary)"),
  ascvd_wimd |> dplyr::count(wimd_f, name = "N") |> dplyr::mutate(Cohort = "ASCVD (secondary)")
)
rt10 <- wimd_sum |>
  dplyr::left_join(n_wimd, by = c("Cohort","wimd_f")) |>
  dplyr::select(Cohort, WIMD_quintile = wimd_f, N,
                `LDL at target (%)` = pct_ldl_target,
                `SBP at target (%)` = pct_sbp_target,
                `On statin (%)` = pct_on_statin,
                `On AHT (%)` = pct_on_aht) |>
  dplyr::mutate(dplyr::across(where(is.numeric), \(x) round(x, 1))) |>
  .mask_all_n()
kbl(rt10) |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```


<br>

### RT11. Supplementary Figure 2 — Treatment receipt by risk quartile (with N per quartile)
```{r rt11_riskq, echo=FALSE}
n_rq <- dplyr::bind_rows(
  dm_q    |> dplyr::filter(!is.na(risk_q)) |>
    dplyr::count(risk_q, name = "N") |> dplyr::mutate(Cohort = "DM (primary)"),
  ascvd_q |> dplyr::filter(!is.na(risk_q)) |>
    dplyr::count(risk_q, name = "N") |> dplyr::mutate(Cohort = "ASCVD (secondary)")
)
rt11 <- rq_sum |>
  dplyr::left_join(n_rq, by = c("Cohort","risk_q")) |>
  dplyr::select(Cohort, Risk_quartile = risk_q, N,
                `LDL at target (%)` = pct_ldl_target,
                `SBP at target (%)` = pct_sbp_target,
                `On statin (%)` = pct_on_statin,
                `On AHT (%)` = pct_on_aht) |>
  dplyr::mutate(dplyr::across(where(is.numeric), \(x) round(x, 1))) |>
  .mask_all_n()
kbl(rt11) |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```


<br>

### RT12. Supplementary Figure 3 — Treatment receipt by sex (with N per group)
```{r rt12_sex, echo=FALSE}
n_sex <- dplyr::bind_rows(
  dm_sex    |> dplyr::count(sex_f, name = "N") |> dplyr::mutate(Cohort = "DM (primary)"),
  ascvd_sex |> dplyr::count(sex_f, name = "N") |> dplyr::mutate(Cohort = "ASCVD (secondary)")
)
rt12 <- sex_sum |>
  dplyr::left_join(n_sex, by = c("Cohort","sex_f")) |>
  dplyr::select(Cohort, Sex = sex_f, N,
                `LDL at target (%)` = pct_ldl_target,
                `SBP at target (%)` = pct_sbp_target,
                `On statin (%)` = pct_on_statin,
                `On AHT (%)` = pct_on_aht) |>
  dplyr::mutate(dplyr::across(where(is.numeric), \(x) round(x, 1))) |>
  .mask_all_n()
kbl(rt12) |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```


<br>

### RT13. Supplementary Figures 4–5 and Figure 5 — Treatment gap at WIMD × risk quartile intersection (with N per cell)
```{r rt13_interact, echo=FALSE}
n_ix <- dplyr::bind_rows(
  ix_dm    |> dplyr::filter(!is.na(risk_q)) |>
    dplyr::count(wimd_f, risk_q, name = "N") |> dplyr::mutate(Cohort = "DM (primary)"),
  ix_ascvd |> dplyr::filter(!is.na(risk_q)) |>
    dplyr::count(wimd_f, risk_q, name = "N") |> dplyr::mutate(Cohort = "ASCVD (secondary)")
)
rt13 <- ix_sum |>
  dplyr::left_join(n_ix, by = c("Cohort","wimd_f","risk_q")) |>
  dplyr::select(Cohort, WIMD_quintile = wimd_f, Risk_quartile = risk_q, N,
                `LDL at target (%)` = pct_ldl_target,
                `On statin (%)` = pct_on_statin,
                `SBP at target (%)` = pct_sbp_target,
                `On AHT (%)` = pct_on_aht,
                `LDL untreated (%)` = pct_ldl_untreated,
                `LDL undertreated (%)` = pct_ldl_undertreated,
                `SBP untreated (%)` = pct_sbp_untreated,
                `SBP undertreated (%)` = pct_sbp_undertreated) |>
  dplyr::mutate(dplyr::across(where(is.numeric), \(x) round(x, 1))) |>
  dplyr::arrange(Cohort, WIMD_quintile, Risk_quartile) |>
  .mask_all_n()
kbl(rt13) |>
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```

